### 2021-06-29 01-preliminary.Rmd
getting wordbank data

---
title: "Looking for CDI Bias"
author: "Nathan & George"
date: "6/29/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) # requires package loading, not just referencing. 
library(mirtCAT)
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English.

```{r get-wordbank-data, echo=T}
get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

table(d_demo$age)
table(d_demo$mom_ed)

sum(is.na(d_demo$mom_ed)) # many missing

```

We have data from `r nrow(d_demo)` children for the WG form.
<!-- writing reminder -->

```{r get-wordbank-items}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(produces = as.numeric(value == "produces")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

d_mat <- d_long_wg %>% select(data_id, num_item_id, produces) %>% # ToDo: filter out non-word items here
  pivot_wider(id_cols = data_id, names_from = num_item_id, values_from = produces)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word")

table(en_wg$category)
```

# DIF Analysis with mirt

```{r}
# https://rdrr.io/cran/mirt/man/DIF.html

mirt(d_mat, 1, "2PL")

# eventually: DIF(multigroupmodel based on sex or mom_ed)
```