### 2021-07-02 01-preliminary.Rmd
fitting basic 2PL model to WG production

---
title: "Looking for CDI Bias"
author: "Nathan, George, and Mike"
date: "`Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English. We'll start with production data from the CDI: Words & Gestures (WG) form.
We'll look for DIF based on 1) sex and 2) SES (high vs. low). 
Questions: How do we choose a dozen anchor items (that we expect to be unbiased)?

```{r get-wordbank-data, eval=F, echo=F}
# Run Once
#get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

# 8-18 month-olds
table(d_demo$age)

# mother's years of education = proxy for SES
table(d_demo$mom_ed)
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

#sum(is.na(d_demo$mom_ed)) # many mom_ed missing

```

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(produces = as.numeric(value == "produces")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word") %>% select(-complexity_category, -item_id)

# subjects not producing any words yet (can't fit model with them)
bad_Ss = d_demo[which(d_demo$production==0),]$data_id

d_long_wg <- d_long_wg %>% left_join(en_wg %>% select(num_item_id, definition)) %>%
  filter(!is.na(definition))

d_mat <- d_long_wg %>% select(data_id, definition, produces) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = produces)

save(d_mat, d_demo, en_wg, file=here("data/en_wg.Rdata"))
```

```{r load-data}
# load saved CDI:WG data
load(here("data/en_wg.Rdata")) # d_demo, d_mat, en_wg (items)

d_demo <- d_demo %>% filter(production!=0) %>%
  arrange(data_id)

# there are more IDs in d_mat than in d_demo; remove those
d_mat <- d_mat %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)

sids = d_mat$data_id
d_mat$data_id = NULL
d_mat = data.matrix(d_mat)
row.names(d_mat) = sids

table(en_wg$category)
```

We have data from `r nrow(d_demo)` children for the WG form.

## Fit baseline 2PL model

```{r}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=1000))

coefs_2pl <- as_tibble(coef(m1, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m1, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat), 
                             ability = fscores(m1, method = "MAP")[,1])

save(file = here("data/eng_wg_mod_2pl.Rds"), "m1", "coefs_2pl") # "fscores_2pl",
```

# DIF Model: Sex

```{r}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

sex_group = d_demo$sex

mg1 <- multipleGroup(dat, model = 1, group = sex_group, verbose = FALSE) 
mg2 <- multipleGroup(dat, model = 1, group = sex_group, 
                     invariance = 'slopes', verbose = FALSE)

anova(mg2, mg1)
```