### 2021-07-06 01-preliminary.Rmd
looking at item DIF

---
title: "Looking for CDI Bias"
author: "Nathan, George, and Mike"
date: "`Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English. We'll start with production data from the CDI: Words & Gestures (WG) form.
We'll look for DIF based on 1) sex and 2) SES (high vs. low). 
Questions: How do we choose a dozen anchor items (that we expect to be unbiased)?

```{r get-wordbank-data, eval=F, echo=F}
# Run Once
#get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

# 8-18 month-olds
table(d_demo$age)

# mother's years of education = proxy for SES
table(d_demo$mom_ed)
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

#sum(is.na(d_demo$mom_ed)) # many mom_ed missing

```

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(produces = as.numeric(value == "produces")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word") %>% select(-complexity_category, -item_id)

# subjects not producing any words yet (can't fit model with them)
bad_Ss = d_demo[which(d_demo$production==0),]$data_id

d_long_wg <- d_long_wg %>% left_join(en_wg %>% select(num_item_id, definition)) %>%
  filter(!is.na(definition))

d_mat <- d_long_wg %>% select(data_id, definition, produces) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = produces)

save(d_mat, d_demo, en_wg, file=here("data/en_wg.Rdata"))
```

```{r load-data}
# load saved CDI:WG data
load(here("data/en_wg.Rdata")) # d_demo, d_mat, en_wg (items)

d_demo <- d_demo %>% 
  filter(production!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id)

# there are more IDs in d_mat than in d_demo; remove those
d_mat <- d_mat %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)

sids = d_mat$data_id
d_mat$data_id = NULL
d_mat = data.matrix(d_mat)
row.names(d_mat) = sids

table(en_wg$category)
```

We have data from `r nrow(d_demo)` children for the WG form.

## Fit baseline 2PL model

```{r}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=500))

coefs_2pl <- as_tibble(coef(m1, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m1, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat), 
                             ability = fscores(m1, method = "MAP")[,1])

save(file = here("data/eng_wg_mod_2pl.Rds"), "m1", "coefs_2pl") # "fscores_2pl",
```

# DIF Model: Sex

```{r}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

sex_group = as.character(d_demo$sex)

# ToDo: define ses_group ("high" for mom_ed > X, "low") (choose X ..high school?)

mg1 <- multipleGroup(d_mat, 1, group = sex_group, SE = T) 
# Error: Multiple Group model will not be identified without proper constraints (groups contain missing data patterns where item responses have been completely omitted or, alternatively, the number of categories within each group is not equal to the total number of categories)
mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=1000), 
                     invariance = 'slopes', verbose = T)

DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group,
                             invariance=c('slopes', 'intercepts', 'free_var','free_means'))
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group,  #fixed means
                             invariance=c('slopes', 'intercepts', 'free_var'))
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group,
                             invariance=c('slopes', 'intercepts'))

extract.mirt(mod_fullconstrain, 'time') #time of estimation components

residuals(mod_sc1)

save(file = here("data/eng_wg_sex_slopes_invariant.Rds"), mg2)

Mit = as_tibble(coef(mg2, simplify=T)$Male$items) %>%
  mutate(definition = rownames(coef(mg2, simplify=T)$Male$items)) %>%
  select(-g, -u) %>% 
  rename(d_m = d)
Fit = as_tibble(coef(mg2, simplify=T)$Female$items) %>%
  mutate(definition = rownames(coef(mg2, simplify=T)$Female$items)) %>%
  select(-g, -u) %>%
  rename(d_f = d)

combo <- Mit %>% left_join(Fit) %>%
  mutate(d_diff = d_f - d_m,
         d_diff_abs = abs(d_diff))

thresh = median(combo$d_diff_abs) + sd(combo$d_diff_abs)

anova(m1, mg2) # AIC prefers group model, BIC prefers 2PL
mg2s <- summary(mg2)
fem = mg2s[[1]]$rotF
mal = mg2s[[2]]$rotF

plot(x=fem, y=mal) # slopes are constrained to be the same...
# where are the difficulties? 
```