### 2021-07-30 02-preliminary-WGcomp.Rmd
multipleGroup model comparisons

---
title: "Looking for CDI Bias: WG Comprehension"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
library(GGally)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English. We'll start with production data from the CDI: Words & Gestures (WG) form.
We'll look for DIF based on 1) sex and 2) SES (high vs. low). 
Questions: How do we choose a dozen anchor items (that we expect to be unbiased)?

```{r get-wordbank-data, eval=F, echo=F}
# Run Once
#get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

# 8-18 month-olds
table(d_demo$age)

# mother's years of education = proxy for SES
table(d_demo$mom_ed) %>% kable()
# skewed: only 72 with Primary/Some Secondary (most have Some College +)

#sum(is.na(d_demo$mom_ed)) # many mom_ed missing

```

Let's do comprehension data.

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(produces = as.numeric(value == "yes" | value == "produces")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word") %>% select(-complexity_category, -item_id)

# subjects not producing any words yet (can't fit model with them)
bad_Ss = d_demo[which(d_demo$comprehension==0),]$data_id # 8 bad comprehension Ss

d_long_wg <- d_long_wg %>% left_join(en_wg %>% select(num_item_id, definition)) %>%
  filter(!is.na(definition))

d_mat <- d_long_wg %>% select(data_id, definition, produces) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = produces)

save(d_mat, d_demo, en_wg, file=here("data/en_wg_comp.Rdata"))
```

```{r load-data}
# load saved CDI:WG data
load(here("data/en_wg_comp.Rdata")) # d_demo, d_mat, en_wg (items)

d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id)

# there are more IDs in d_mat than in d_demo; remove those
d_mat <- d_mat %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)

sids = d_mat$data_id
d_mat$data_id = NULL
d_mat = data.matrix(d_mat)
row.names(d_mat) = sids

table(d_demo$mom_ed)
# high SES = mom_ed >= Some College (739 high, 324 low-SES)
d_demo <- d_demo %>% mutate(ses_group = ifelse(as.numeric(mom_ed)>4, "high", "low"))

table(en_wg$category) %>% kable()
```

We will fit comprehension data from `r nrow(d_demo)` children for the WG form.

## Fit baseline 2PL model

```{r, eval=F}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=1000))

coefs_2pl <- as_tibble(coef(m1, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m1, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat), 
                             ability = fscores(m1, method = "MAP")[,1])

save(file = here("data/eng_wg_comp_mod_2pl.Rds"), "m1", "coefs_2pl", "fscores_2pl") # ,
```

# DIF Model: Sex

```{r, eval=F}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

sex_group = as.character(d_demo$sex)

# ToDo: define ses_group ("high" for mom_ed > X, "low") (choose X ..high school?)

mg1 <- multipleGroup(d_mat, 1, group = sex_group) # this example should work, according to docs
mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
                     invariance = c("free_var", "free_means")) 
# Error: Multiple Group model will not be identified without proper constraints (groups contain missing data patterns where item responses have been completely omitted or, alternatively, the number of categories within each group is not equal to the total number of categories)

# unless there are possibly items

mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))

# constrain slopes, freely estimate latent group means (reference group constrained to a vector of 0)
mg3 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes', 'free_means'), verbose = T)

# constrain slopes, allow varying group means and variances -- very slow!
#mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
#                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

# fixed means, varying group variance
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts', 'free_var'))
# fully-constrained
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

# constrain intercepts to be equal across groups, but allow group slopes to vary
m_int <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                     invariance = 'intercepts')

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by AIC/BIC

anova(mg2, mod_fullconstrain) # AIC prefers slope-invariant model; BIC prefers fully-constrained model

anova(m_int, mod_fullconstrain) # AIC prefers intercept-constrained model; BIC prefers fully-constrained

anova(mg3, mg2) # AIC//BIC prefer model with varying group means
#anova(mg3, mg4)

coef(mg3)[[2]]$GroupPars # Female intercept = 0.22 (Male = 0)

#extract.group(m_sc2, "Male")
#extract.group(m_sc2, "Female")

residuals(mod_sc1)

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg2, mg3,  m_sc2, mod_sc1, mod_fullconstrain) # not run: mg4
```

```{r}
load(here("data/eng_wg_comp_sex_sexDIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

extract_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$Male$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Male$items)) %>%
    select(-g, -u) %>% 
    rename(d_m = d)
  Fit = as_tibble(coef(group_model, simplify=T)$Female$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Female$items)) %>%
    select(-g, -u) %>%
    rename(d_f = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_f - d_m,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

# slope-invariant, varying difficulty:
mg2_df = extract_group_df(mg2)
mg3_df = extract_group_df(mg3)

# same item estimates with/without free_means? 
cor(mg2_df$d_f, mg3_df$d_f) # .9999
cor(mg2_df$d_m, mg3_df$d_m) # .9998

# same difficulties:
mod_sc1_df = extract_group_df(mod_sc1)

# fully-constrained is just the same as ordinary 2PL model, no? but we can use this model for DIF
mod_fullconstrain_df = extract_group_df(mod_fullconstrain)
# DIF(mod_fullconstrain, which.par = 'd')

#mod_int_df = extract_group_df(m_int)

mod_sc2_df = extract_group_df(m_sc2)

#anova(m1, mg2) # AIC prefers group model, BIC prefers 2PL
```

### Items with a large difference in difficulty across sex

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg3_df$d_diff_abs) + sd(mg3_df$d_diff_abs)
big_dif = mg3_df[which(mg3_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
#fem = mg2s[[1]]$rotF
#mal = mg2s[[2]]$rotF
big_dif %>% relocate(definition) %>%
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Plot item ease for males vs. females

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

## Ability distribution comparison

What are the ability distributions by sex in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=sex)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg3 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg3, method = "MAP")[,1])

fscores_mg2 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg2, method = "MAP")[,1])

# very similar
#plot(fscores_mg3$ability_sex, fscores_mg2$ability_sex, pch='.')

d_demo <- d_demo %>% left_join(fscores_mg2 %>% 
                                 mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo, columns = c("age","comprehension","ability","ability_sex"),
        ggplot2::aes(colour=sex, alpha=.5)) + theme_classic()
```

```{r}
d_demo %>% group_by(sex) %>% 
  summarise(ability_2pl = mean(ability),
            ability_sex = mean(ability_sex)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```

Average male ability is slightly higher in the sex group model (i.e., there is a larger female advantage in the standard 2PL model), showing that the 36 items that are easier for females boost female ability estimates.

## DIF Model: SES

```{r, eval=F}
ses_group = as.character(d_demo$ses_group)

#mg1 <- multipleGroup(d_mat, 1, group = ses_group, SE = T) 
mg2 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)

mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg2, mg3) # AIC/BIC prefer model with varying group mean

#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))
mod_sc1 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000), #fixed means
                             invariance=c('slopes', 'intercepts', 'free_var'))
mod_fullconstrain <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

#extract.group(m_sc2, "high")
#extract.group(m_sc2, "low")

residuals(mod_sc1)

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain)
```

```{r}
load(here("data/eng_wg_comp_SES_DIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

extract_ses_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$high$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$high$items)) %>%
    select(-g, -u) %>% 
    rename(d_hi = d)
  Fit = as_tibble(coef(group_model, simplify=T)$low$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$low$items)) %>%
    select(-g, -u) %>%
    rename(d_lo = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_hi - d_lo,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

mg2_df = extract_ses_group_df(mg2)
mg3_df = extract_ses_group_df(mg3)

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .24
```

### Items with a large difference in difficulty across SES

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg3_df$d_diff_abs) + sd(mg3_df$d_diff_abs)
big_dif = mg3_df[which(mg3_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
big_dif %>% relocate(definition) %>% 
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Plot item ease for high vs. low SES

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg3_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

## Ability distribution comparison

What are the ability distributions by SES in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=ses_group)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg2 <- tibble(data_id = rownames(d_mat[which(!is.na(d_demo$ses_group)),]), 
                             ability_ses = fscores(mg2, method = "MAP")[,1])

d_demo_ses <- d_demo %>% 
  filter(!is.na(ses_group)) %>% 
  left_join(fscores_mg2 %>% mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo_ses, columns = c("age","comprehension","ability","ability_ses"),
        ggplot2::aes(colour=ses_group, alpha=.5)) + theme_classic()
```

```{r, echo=F}
d_demo_ses %>% group_by(ses_group) %>% 
  summarise(ability_2pl = mean(ability),
            ability_ses = mean(ability_ses)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```