### 2021-08-17 ws_prod_report.Rmd
outlined report

---
title: "CDI:WS Production Bias Report"
author: "Nathan, George, & Mike"
date: "8/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(tidyverse)
library(difR)
library(mirt)
library(janitor)
library(here)
require(kableExtra)

# created in GLIMMER_production.R
load(here("data/en_ws_production.Rdata")) # d_demo, d_mat, en_wg (items)
source("report_helper_functions.R")
```

# Introduction

## Item-Response Theory

## Differential Item Function

### Dataset

We analyze Wordbank CDI Words & Sentences form data from `r nrow(d_demo)` children aged `r min(d_demo$age)` to `r max(d_demo$age)`.

```{r, sex-table}
#d_demo <- d_demo %>% mutate(relevel() # remove empty "Other" level?

table(d_demo$sex) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

## Goals

For each axis (sex, SES, race/ethnicity)
1) show GLIMMER plots (small sample of words)
2) histograms of DIF: fit normal distro, exclude >2SD (list those items)
  - overall statistic measuring DIF of the test? sd(normal), but that doesn't account for overall group difference (subtract mean?) (DTF definition/statistic)
3) re-fit IRT model, show new GLIMMER (discuss: does the test look much more fair, or are there new outliers?)
4) suggested new words from CHILDES (equal frequency for boys/girls) [but we can't do this for SES/race unless CHILDES has demo data]

## Sex DIF

### GLIMMER plots

```{r show-sex-glimmer, fig.width=5, fig.height=5.5, caption="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
load("data/glimmer_prodWS_models.Rds")

plot_glimmer(mod_intuitive_sex, colnames(d_mat)[2:ncol(d_mat)])

```

### Removing Biased Items

```{r}
mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
sex_hist <- item_difficulty_difference_histogram(mm_sex)

orig_sex <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06 (mean + 2SD)
sex_hist
```

```{r sex-diff-table}
items_to_remove <- orig_sex %>% filter(extreme==T)

items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:30%;'")
```

ToDo: filter items_to_remove out of d_mat, re-run 1PL IRT model

see what the new distribution of item difficulty differences looks like (are there any more outliers?)

```{r, eval=F}
# chunk to run once:
mod_intuitive_sex <- fit_mod_intuitive(d_mat, sex_group)

save()
```

```{r}
load()
```

## SES DIF

```{r ses-table}

```

```{r}
plot_glimmer(mod_intuitive_ses, colnames(d_mat), plotName="GLIMMER_ses_prodWS")

```

## Race/Ethnicity DIF

```{r, race-table}
# breakdown white/non-white
```

# Discussion

Where to go from here -- other languages, regional variation, 
- using adult word vectors to find unbiased items?