### 2021-08-25 03-preliminary-WSprod.Rmd
added WS production fits

---
title: "Looking for CDI Bias: WS Production"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
library(GGally)
library(ggpubr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English. We'll now turn to production data from the CDI: Words & Sentences (WS) form.
We'll look for DIF based on 1) sex, 2) SES (high vs. low), and 3) race/ethnicity.

```{r load-data}
# see GLIMMER_production.R
load(here("data/en_ws_production.Rdata"))

d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))

# there are more IDs in d_mat than in d_demo; remove those
d_mat <- d_mat %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)

# only 4090 kids with demographic variables

sids = d_mat$data_id
d_mat$data_id = NULL
d_mat = data.matrix(d_mat)
row.names(d_mat) = sids

sex_group = as.character(d_demo$sex)
ses_group = as.character(d_demo$ses_group)
eth_group = as.character(d_demo$eth_group)

table(d_demo$mom_ed)

table(en_ws$category) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We will fit comprehension data from `r nrow(d_demo)` children for the WS form.

## Fit baseline 2PL model

```{r, eval=F}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=1000))

coefs_2pl <- as_tibble(coef(m1, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m1, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat), 
                             ability = fscores(m1, method = "MAP")[,1])

sex_group = as.character(d_demo$sex)

save(file = here("data/eng_ws_prod_mod_2pl.Rds"), "m1", "coefs_2pl", "fscores_2pl") 
```

## GLIMMER Plots

(See external PDFs and GLIMMER_WGcomp.R)

# DIF Model: Sex

```{r, eval=F}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

# won't work if some items are all 0 (or 1?) for one sex
mg1 <- multipleGroup(d_mat, 1, group = sex_group,
                     technical=list(NCYCLES=4000)) 
#mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
#                     invariance = c("free_var", "free_means"))

mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))

# constrain slopes, freely estimate latent group means (reference group constrained to a vector of 0)
mg3 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes', 'free_means'), verbose = T)

# constrain slopes, allow varying group means and variances -- very slow!
mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

# fixed means, varying group variance
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts', 'free_var'))
# fully-constrained
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

# constrain intercepts to be equal across groups, but allow group slopes to vary
m_int <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                     invariance = 'intercepts')

anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
anova(mg3, mg4) # AIC/BIC prefer mg3, with varying group means but equal variance

# Takehome: by model comparison, we find that equal slopes (across sex) and equal variance, but varying group means (and item difficulties) are justified.

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc1 by AIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc2 by AIC

anova(mod_fullconstrain, mg2) # AIC and BIC both prefer slope-invariant model

anova(m_int, mod_fullconstrain) # AIC prefers intercept-constrained model; BIC prefers fully-constrained

coef(mg3)[[2]]$GroupPars # Female intercept = 0.047 (Male = 0)

#extract.group(m_sc2, "Male")
#extract.group(m_sc2, "Female")

residuals(mod_sc1)

mirtCluster()
mg3_dif = DIF(mg3, c('d'))

save(file = here("data/eng_ws_prod_sex_sexDIF.Rds"), 
     mg3_dif, mg1, mg2, mg3, mg4, m_sc2, mod_sc1, mod_fullconstrain) 
```

```{r}
load(here("data/eng_ws_prod_sex_sexDIF.Rds"))
load(here("data/eng_ws_prod_mod_2pl.Rds"))

extract_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$Male$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Male$items)) %>%
    select(-g, -u) %>% 
    rename(d_m = d)
  Fit = as_tibble(coef(group_model, simplify=T)$Female$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Female$items)) %>%
    select(-g, -u) %>%
    rename(d_f = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_f - d_m,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

# slope-invariant, varying difficulty:
mg2_df = extract_group_df(mg2)
mg3_df = extract_group_df(mg3)

# same item estimates with/without free_means? 
#cor(mg2_df$d_f, mg3_df$d_f) # .999
#cor(mg2_df$d_m, mg3_df$d_m) # .999

#anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
#anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
```

### Items with DIF in difficulty

`r nrow(subset(mg3_dif, p<.05))` items show DIF based on a $\chi^2$ test yielding $p<.05$.
Below we show the `r nrow(subset(mg3_dif, p<.001))` items with $p<.001$.

```{r}
mg3_dif %>% arrange(BIC) %>% filter(p < .001) %>%
  select(-AICc, -SABIC, -HQ, -df) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Items with a large difference in difficulty across sex

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg3_df$d_diff_abs) + sd(mg3_df$d_diff_abs)
big_dif = mg3_df[which(mg3_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
#fem = mg2s[[1]]$rotF
#mal = mg2s[[2]]$rotF
big_dif %>% relocate(definition) %>%
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Plot item ease for males vs. females

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant free group means model (mg3).

```{r}
sex_anchors = mg3_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_m, .after=d_f)

sex_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```

## Ability distribution comparison

What are the ability distributions by sex in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=sex)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg3 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg3, method = "MAP")[,1])

fscores_mg2 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg2, method = "MAP")[,1])

# very similar
#plot(fscores_mg3$ability_sex, fscores_mg2$ability_sex, pch='.')

d_demo <- d_demo %>% left_join(fscores_mg3 %>% 
                                 mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo, columns = c("age","comprehension","ability","ability_sex"),
        ggplot2::aes(colour=sex, alpha=.5)) + theme_classic()
```

```{r}
d_demo %>% group_by(sex) %>% 
  summarise(ability_2pl = mean(ability),
            ability_sex = mean(ability_sex)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```

Average male ability is slightly higher in the sex group model (i.e., there is a larger female advantage in the standard 2PL model), showing that the 36 items that are easier for females boost female ability estimates.

Extracting the estimated ability intercept for males from mg3: `r round(coef(mg3)[["Male"]]$GroupPars[1],2)`. 
(Why is this positive? Females should have higher ability (see plot).)

## SES-related DIF

Now we will look for SES-based DIF, using mother's education as a proxy for SES. 
Here's the distribution of mom_ed in our data:

```{r, mom-ed-tab, echo=F}
table(d_demo$mom_ed) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We (arbitrarily) define high SES as their mother having at least some college (N=`r length(which(d_demo$ses_group=="high"))`), and low SES as having at most secondary education (N=`r length(which(d_demo$ses_group=="low"))`).

```{r, eval=F}
ses_group = as.character(d_demo$ses_group)

mg1 <- multipleGroup(d_mat, 1, group = ses_group, SE = T, technical=list(NCYCLES=4000)) 
mg2 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg1, mg2) # AIC/BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer slope-invariant model without varying group mean

#DIF(mg1, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))
mod_sc1 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000), #fixed means
                             invariance=c('slopes', 'intercepts', 'free_var'))
mod_fullconstrain <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

#extract.group(m_sc2, "high")
#extract.group(m_sc2, "low")

#residuals(mod_sc1)

mg2_dif = DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.
#mg3_dif = DIF(mg3, c('d')) # (though varying group mean not justified according to AIC/BIC)

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg1, mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain,
     mg2_dif) # mg3_dif
```

```{r}
load(here("data/eng_wg_comp_SES_DIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

extract_ses_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$high$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$high$items)) %>%
    select(-g, -u) %>% 
    rename(d_hi = d)
  Fit = as_tibble(coef(group_model, simplify=T)$low$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$low$items)) %>%
    select(-g, -u) %>%
    rename(d_lo = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_hi - d_lo,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

mg2_df = extract_ses_group_df(mg2)
mg3_df = extract_ses_group_df(mg3)

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .25
```

### Items with DIF in difficulty

`r nrow(subset(mg3_dif, p<.05))` items show DIF based on a $\chi^2$ test yielding $p<.05$.
We show these items below.

```{r}
mg3_dif %>% arrange(BIC) %>% filter(p < .05) %>%
  select(-AICc, -SABIC, -HQ, -df) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Items with a large difference in difficulty across SES

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg2_df$d_diff_abs) + sd(mg2_df$d_diff_abs)
big_dif = mg2_df[which(mg2_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
big_dif %>% relocate(definition) %>% 
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```

### Plot item ease for high vs. low SES

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant model (mg2).

```{r ses-anchors, echo=F}
ses_anchors = mg2_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_hi, .after=a1)

ses_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```

## Ability distribution comparison

What are the ability distributions by SES in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=ses_group)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg2 <- tibble(data_id = rownames(d_mat[which(!is.na(d_demo$ses_group)),]), 
                             ability_ses = fscores(mg2, method = "MAP")[,1])

d_demo_ses <- d_demo %>% 
  filter(!is.na(ses_group)) %>% 
  left_join(fscores_mg2 %>% mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo_ses, columns = c("age","comprehension","ability","ability_ses"),
        ggplot2::aes(colour=ses_group, alpha=.5)) + theme_classic()
```

```{r, echo=F}
d_demo_ses %>% group_by(ses_group) %>% 
  summarise(ability_2pl = mean(ability),
            ability_ses = mean(ability_ses)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```