### 2021-09-09 word_freq.Rmd
smoothed counts and Childes words

---
title: "Word_freq"
author: "Nathan, George, Mike"
date: "8/27/2021"
output: html_document
---

```{r setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(dbplyr)
library(wordbankr)
library(childesr)
library(glue)
source("report_helper_functions.R")
```

```{r get-data, include=FALSE}
d_eng_na <- get_transcripts(collection = "Eng-NA")
head(d_eng_na)
sex_range <- get_participants(sex = "male", role_exclude = "Target_Child")
```

```{r wordbank}
data_mode <- "remote"
languages <- "English (American)"
admins <- get_administration_data(mode = data_mode) %>%
  select(language, form, sex, age, data_id) %>%
  filter(language %in% languages)
words <- get_item_data(mode = data_mode) %>%
  filter(type == "word", language %in% languages) %>%
  select(language, form, lexical_class, category, uni_lemma, definition,
         item_id) %>%
  mutate(num_item_id = as.numeric(substr(item_id, 6, nchar(item_id))),
         definition = tolower(definition))
```

```{r counts, include=FALSE}
#only english
exc_target_types <- get_types(#target_child = "Adam",
                          role_exclude = "target_child",
                          language = "eng")
                          # type = c("dog", "ball"))

sex_group <- as.character(exc_target_types$target_child_sex)

#have to nest because code expects nested DF
#uni_lemma and definition are basically the same in english 
items <- get_item_data(language = "English (American)", form = "WS")
items_nested <- items %>%
  filter(type == "word", !str_detect(definition, "name")) %>%
  mutate(uni_lemma = definition) %>%
  nest(items = -c(language, uni_lemma))
uni_lemma_map <- build_uni_lemma_map(items_nested)

# target_types <- exc_target_types[exc_target_types$gloss %in% CDI_words, ]

#groups by sex and words(lower case) so now theres individual words with total count per gender. 
target_type_counts <- exc_target_types %>%
  filter(!is.na(target_child_sex)) %>%
  mutate(gloss = str_to_lower(gloss)) %>%
  group_by(language, gloss, target_child_sex) %>%
  summarise(count = sum(count))
#stem childes words and get counts to use as recomended replacements

#stem the gloss(the words in Childes)
type_counts_stemmed <- target_type_counts %>%
  mutate(gloss_stem = stem(gloss, convert_lang_stemmer("English (American)")))

#group stemmed words and counts
childes_replace <- type_counts_stemmed %>%
  group_by(gloss_stem, target_child_sex) %>%
  summarise(count = sum(count))

normalized_counts <- childes_replace %>%
  group_by(target_child_sex) %>%
  mutate(prob = count/sum(count), 
         count_per_mil =  prob * 1000000)

#totals of counts
total_counts <- childes_replace %>%
   group_by(target_child_sex) %>%
   summarize(count = sum(count))

#joins the uni_lemmas from CDI with both gloss and stemmed gloss in Childes to get only the overlapping words
type_counts_mapped <- type_counts_stemmed %>%
  ungroup() %>%
  select(-language) %>%
  inner_join(uni_lemma_map %>% mutate(option = tolower(option)) %>% rename(gloss = option)) %>%
  inner_join(uni_lemma_map %>% mutate(option = tolower(option)) %>% rename(gloss_stem = option)) %>%
  select(-language, -gloss_stem)

# TODO: normalize counts (over CDI words or over all of CHILDES? per sex or joined?)
#re-summarizes by counts 
type_counts <- type_counts_mapped %>%
  group_by(target_child_sex, uni_lemma) %>%
  summarise(count = sum(count))
  #1 is female 2 is male
  #mutate(count_per_mil = 1000000 * (count/ifelse(target_child_sex == "male", total_counts$count[2], total_counts$count[1])))

#finds missing items and makes their counts = 1
#missing mostly stared words like "penis*" and "church*"
missing_items <- items_nested %>%
  filter(!(uni_lemma %in% type_counts$uni_lemma)) %>%
  select(uni_lemma) %>%
  mutate(female = 1, male = 1)

#adds back in the unfound words with count 1 and adds ones to all other words
type_counts_smoothed <- type_counts %>%
  mutate(count = count + 1) %>%
  pivot_wider(names_from = target_child_sex, values_from = count, values_fill = 1) %>%
  bind_rows(missing_items) %>%
  pivot_longer(cols = c(female, male), names_to = "sex", values_to = "count") %>%
  #1 is female 2 is male
  mutate(count_per_mil = 1000000 * (count/ifelse(sex == "male", total_counts$count[2], total_counts$count[1])))

# TODO: normalize counts (over CDI words or over all of CHILDES? per sex or joined?)
#smoothing
type_freqs <- type_counts_smoothed %>%
  mutate(freq = log(count / sum(count)))

# saveRDS(type_counts, "data/type_counts.rds")
# saveRDS(type_freqs, "data/type_freqs.rds")
#saveRDS(normalized_counts, "data/normalized_childes_counts.rds")
#saveRDS(type_counts_smoothed, "data/type_counts_smoothed.rds")

# TODO: this is log ratio of probabilities, if doing log odds ratio transform to odds before taking difference
type_diffs <- type_freqs %>%
  select(-count) %>%
  pivot_wider(names_from = sex, values_from = freq) %>%
  mutate(freq_ratio = female - male)

qplot(type_diffs$freq_ratio)

CDI_words <- c(items$definition)
#potential issue in that the words they count include "accidentally" and "called" rather than the general form so we may need to lemmatize all the words in both or something. plus church* and mommy*
#plus id 120 is the name nathan for the child nathan which is child's_name in CDI but isn't reported as such in the wordbank and "drink (action)" is not recorded as such. 
#target_types <- exc_target_types[exc_target_types$gloss %in% CDI_words, ] 
#sum the counts for spoken to boys removing the na sex values. 
#note truck had a huge difference in favor of boys and purse had a very small difference for girls so I wonder if this is perfect
word_total <- sum(target_types[target_types$gloss == "nurse",]$count, na.rm=TRUE)
boy_tmw <- sum(target_types[(target_types$target_child_sex == "male")&(target_types$gloss == "nurse"),]$count, na.rm=TRUE)
girl_tmw <- sum(target_types[(target_types$target_child_sex == "female")&(target_types$gloss == "nurse"),]$count, na.rm=TRUE)
#freq_dif returns 
num <- freq_dif(target_types, "truck")
```

```{r find words, include=FALSE}
best_words = c()
for(word in CDI_words) {
  if((word %in% target_types$gloss)& (abs(freq_dif(target_types, word)) < 10) ){
    #print(abs(freq_dif(target_types, word)))
    best_words <- append(best_words, word)
  }
}
#finds no words that have a difference of 0 exactly 
```