### 2021-09-21 ws_prod_report.Rmd
add comparison of by-sex CHILDES word freqs to DIF

---
title: "CDI:WS Production Bias Report"
author: "Nathan, George, & Mike"
date: "8/17/2021"
output: html_document
fig_width: 3
fig_height: 3
fig_align: "center"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(tidyverse)
library(difR)
library(mirt)
library(janitor)
library(here)
require(kableExtra)

library(wordbankr)
library(ggrepel)
library(gridExtra)
library(printr)

# created in GLIMMER_production.R
load(here("data/en_ws_production.Rdata")) # d_demo, d_mat, en_wg (items)
source("report_helper_functions.R")
```

# Introduction

## Item-Response Theory

## Differential Item Function

### Dataset

We analyze Wordbank CDI Words & Sentences form data from `r nrow(d_demo)` children aged `r min(d_demo$age)` to `r max(d_demo$age)`.

```{r, sex-table}
#d_demo <- d_demo %>% mutate(relevel() # remove empty "Other" level?
#need this so that d_demo extists in the right form with you knit.
d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))

table(d_demo$sex) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

## Goals

For each axis (sex, SES, race/ethnicity)
1) show GLIMMER plots (small sample of words)
2) histograms of DIF: fit normal distro, exclude >2SD (list those items)
  - overall statistic measuring DIF of the test? sd(normal), but that doesn't account for overall group difference (subtract mean?) (DTF definition/statistic)
3) re-fit IRT model, show new GLIMMER (discuss: does the test look much more fair, or are there new outliers?)
4) suggested new words from CHILDES (equal frequency for boys/girls) [but we can't do this for SES/race unless CHILDES has demo data]

```{r load data}
load("data/glimmer_prodWS_models.Rds")
load(here("data/glimmer_prodWS_model_noExt.Rds")) # add this in once I've saved all the new models
```

## Sex DIF

### GLIMMER plots

```{r show-sex-glimmer,echo= FALSE, warning=FALSE,message=FALSE,error=FALSE, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}

plot_glimmer(mod_intuitive_sex, colnames(d_mat)[2:ncol(d_mat)])

```

### Removing Biased Items
Histogram of Sex Items from GLIMMER
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
sex_hist <- item_difficulty_difference_histogram(mm_sex)
#added the option to change SD threshold
orig_sex <- get_extreme_item_difficulty_differences(mm_sex, 2.25) # min: -.53 max: 1.06 (mean + 2SD) mean: .265
sex_hist
```

```{r sex-diff-table, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
sex_items_to_remove <- orig_sex %>% filter(extreme==T)

sex_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(sex_bias = case_when(d_diff < 0 ~ "Male", d_diff > 0 ~ "Female")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
List of Sex Items that show DIF outside of the 2SD range.(shows `r nrow(sex_items_to_remove)` items that have DIF)

see what the new distribution of item difficulty differences looks like (are there any more outliers?)

```{r, eval=F}
#may no longer be nesesary now that its up top and I have already fit the models but I left it for now. 
d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))

sex_group = as.character(d_demo$sex)

sex_words <- c(sex_items_to_remove$definition) 
#remove words
d_mat_sex <- d_mat[, ! names(d_mat) %in% sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex <- d_mat_sex %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex$data_id
d_mat_sex$data_id = NULL
d_mat_sex = data.matrix(d_mat_sex)
row.names(d_mat_sex) = sids

#refit model now with main DIF items removed
mod_intuitive_sexn <- fit_mod_intuitive(d_mat_sex, sex_group)

plot_glimmer(mod_intuitive_sexn, colnames(d_mat_sex)[2:ncol(d_mat_sex)])

```

```{r show-new-sex-glimmer,echo= FALSE, warning=FALSE,message=FALSE,error=FALSE, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
sex_words <- c(sex_items_to_remove$definition) 
#remove words
d_mat_sex <- d_mat[, ! names(d_mat) %in% sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex <- d_mat_sex %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex$data_id
d_mat_sex$data_id = NULL
d_mat_sex = data.matrix(d_mat_sex)
row.names(d_mat_sex) = sids

plot_glimmer(mod_intuitive_sexn, colnames(d_mat_sex)[2:ncol(d_mat_sex)])

```

Sex Histogram with the DIF Items removed
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
mm_sexn <- extract_group_df(mod_intuitive_sexn, groups=c("Male","Female"))
sex_histn <- item_difficulty_difference_histogram(mm_sexn)

orig_sexn <- get_extreme_item_difficulty_differences(mm_sexn, 2.25) # min: -.20 max: 0.74 (mean + 2SD) mean: 0.268
sex_histn
```

```{r sex-diff-table-new, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
new_sex_items_to_remove <- orig_sexn %>% filter(extreme==T)

new_sex_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(sex_bias = case_when(d_diff < 0 ~ "Male", d_diff > 0 ~ "Female")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
New Sex Items that show DIF (shows `r nrow(new_sex_items_to_remove)` new items that have DIF)

```{r, eval=F}

new_sex_words <- c(new_sex_items_to_remove$definition) 
all_sex_words <- append(sex_words, new_sex_words)
#remove words
d_mat_sex_new <- d_mat[, ! names(d_mat) %in% all_sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex_new <- d_mat_sex_new %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex_new$data_id
d_mat_sex_new$data_id = NULL
d_mat_sex_new = data.matrix(d_mat_sex_new)
row.names(d_mat_sex_new) = sids

#refit model now with main DIF items removed
mod_intuitive_sexn_n <- fit_mod_intuitive(d_mat_sex_new, sex_group)

plot_glimmer(mod_intuitive_sexn_n, colnames(d_mat_sex_new)[2:ncol(d_mat_sex_new)])

```

```{r show-nnew-sex-glimmer,echo= FALSE, warning=FALSE,message=FALSE,error=FALSE, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
new_sex_words <- c(new_sex_items_to_remove$definition) 
all_sex_words <- append(sex_words, new_sex_words)
#remove words
d_mat_sex_new <- d_mat[, ! names(d_mat) %in% all_sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex_new <- d_mat_sex_new %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex_new$data_id
d_mat_sex_new$data_id = NULL
d_mat_sex_new = data.matrix(d_mat_sex_new)
row.names(d_mat_sex_new) = sids

plot_glimmer(mod_intuitive_sexn_n, colnames(d_mat_sex_new)[2:ncol(d_mat_sex_new)])

```

Sex Histogram with the DIF Items removed
```{r 3rd sex hist, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
mm_sexn_n <- extract_group_df(mod_intuitive_sexn_n, groups=c("Male","Female"))
sex_histn_n <- item_difficulty_difference_histogram(mm_sexn_n)

orig_sexn_n <- get_extreme_item_difficulty_differences(mm_sexn_n, 2.25) # min: -.20 max: 0.74 (mean + 2SD) mean: 0.268
sex_histn_n
```

From seeing this my take away is that there are more female DIF Items total but that the extreme ones are mostly male.

```{r sex-words 3rd, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
nnew_sex_items_to_remove <- orig_sexn_n %>% filter(extreme==T)

nnew_sex_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(sex_bias = case_when(d_diff < 0 ~ "Male", d_diff > 0 ~ "Female")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
`r nrow(nnew_sex_items_to_remove)` item that now show DIF

### Suggesting replacement items

```{r}
type_freqs = readRDS("data/type_freqs.rds")
type_counts = readRDS("data/type_counts.rds")
type_counts_smoothed = readRDS("data/type_counts_smoothed.rds")

type_counts_smoothed <- type_counts_smoothed %>%
  pivot_wider(id_cols=uni_lemma, names_from=sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male),
         FtoM = female / male)

# how biased are the items we suggest removing in the corpus counts?
sex_items_to_remove$definition

tmp <- left_join(orig_sex, type_counts_smoothed, by=c("definition"="uni_lemma"))

tmp %>% #filter(extreme==T) %>%
  ggplot(aes(x=propFemale, y=d_diff, color=extreme)) + 
  geom_point(alpha=.5) + theme_bw() +
  ylab("Female - Male Word Difficulty") +
  xlab("Proportion of Utterances to Females in CHILDES")

subset(tmp, extreme==T & propFemale > .47 & propFemale < .53) # hose & dump

cor.test(tmp$propFemale, tmp$d_diff) # .19, p<.001

# for replacements, want equal-frequency by sex (and high-freq overall) words from CHILDES
normalized_childes_counts = readRDS("data/normalized_childes_counts.rds")

childes_rel_freq <- normalized_childes_counts %>% filter(count_per_mil > 10) %>%
  pivot_wider(id_cols=gloss_stem, names_from=target_child_sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male))
# 8717 words to choose from
hist(childes_rel_freq$propFemale)
mean(childes_rel_freq$propFemale, na.rm=T) # .497

replacements <- childes_rel_freq %>% 
  filter(propFemale < .51, propFemale > .49) %>%
  filter(!is.element(gloss_stem, en_ws$definition)) # 182
```

1. Of the 22 items with extreme sex DIF, only 2 of those fall very close to being equally often spoken to boys and girls ('hose' and 'dump')
2. There are many items on the CDI with small difficulty difference (DIF) by sex

## SES DIF
```{r, ses-table}
#d_demo <- d_demo %>% mutate(relevel() # remove empty "Other" level?

table(d_demo$ses_group) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

```{r show-ses-glimmer, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
plot_glimmer(mod_intuitive_ses, colnames(d_mat)[2:ncol(d_mat)])
```

Histogram from SES GLIMMER
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#used this to find out the that the two groups are 0 1 not high low, not sure why
#groups <- c("0","1")
#group_model <- mod_intuitive_ses
#Mit = as_tibble(coef(group_model, simplify=T)[[groups[1]]]$items) %>%

mm_ses <- extract_group_df(mod_intuitive_ses, groups=c("0","1"))
ses_hist <- item_difficulty_difference_histogram(mm_ses)

orig_ses <- get_extreme_item_difficulty_differences(mm_ses, 2.25) # min:-0.97  max:1.08  (mean + 2SD) mean:0.051
ses_hist
```

```{r ses-diff-table, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
ses_items_to_remove <- orig_ses %>% filter(extreme==T)

ses_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(ses_bias = case_when(d_diff < 0 ~ "Low", d_diff > 0 ~ "High")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
List of SES Items that show DIF outside of the 2SD range. (`r nrow(ses_items_to_remove)` items that show DIF)

```{r, eval=F}
#do we need to have removed NA sex kids for the ses model?
d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))

ses_group = as.character(d_demo$ses_group)

ses_words <- c(ses_items_to_remove$definition) 
#remove words
d_mat_ses <- d_mat[, ! names(d_mat) %in% ses_words, drop = F]
# there are more IDs in d_mat_ses than in d_demo; remove those
d_mat_ses <- d_mat_ses %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_ses$data_id
d_mat_ses$data_id = NULL
d_mat_ses = data.matrix(d_mat_ses)
row.names(d_mat_ses) = sids

#refit model now with main DIF items removed
mod_intuitive_sesn <- fit_mod_intuitive(d_mat_ses, ses_group)

plot_glimmer(mod_intuitive_sesn, colnames(d_mat_ses)[2:ncol(d_mat_ses)])

```
```{r show-new-ses-glimmer, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
ses_words <- c(ses_items_to_remove$definition) 
#remove words
d_mat_ses <- d_mat[, ! names(d_mat) %in% ses_words, drop = F]
# there are more IDs in d_mat_ses than in d_demo; remove those
d_mat_ses <- d_mat_ses %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_ses$data_id
d_mat_ses$data_id = NULL
d_mat_ses = data.matrix(d_mat_ses)
row.names(d_mat_ses) = sids

plot_glimmer(mod_intuitive_sesn, colnames(d_mat_ses)[2:ncol(d_mat_ses)])
```

New Histogram with the DIF Items removed 
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#not sure why the groups are now high and low and not 0 and 1
mm_sesn <- extract_group_df(mod_intuitive_sesn, groups=c("low","high"))
ses_histn <- item_difficulty_difference_histogram(mm_sesn)

orig_sesn <- get_extreme_item_difficulty_differences(mm_sesn, 2.25) # min:-0.97  max:1.08  (mean + 2SD) mean:0.051
ses_histn
```

```{r ses-diff-table-new, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
new_ses_items_to_remove <- orig_sesn %>% filter(extreme==T)

new_ses_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(ses_bias = case_when(d_diff < 0 ~ "Low", d_diff > 0 ~ "High")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
List of new SES Items that show DIF outside of the 2SD range. (shows `r nrow(new_ses_items_to_remove)` new items that have DIF)

## Race/Ethnicity DIF
Race is not a stand alone factor; people of each race fall into every combination of the other axes we have analyzed. Perhaps whiteness and SES are highly correlated in our data. Any binerized set of data will have members of each group that are different across another axis. It will perhap be worth while to specify even further such as the effect of white vs. non-white but only amongst high SES kids.

```{r, eth-table}
#d_demo <- d_demo %>% mutate(relevel() # remove empty "Other" level?

table(d_demo$eth_group) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

```{r show-eth-glimmer, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
plot_glimmer(mod_intuitive_eth, colnames(d_mat)[2:ncol(d_mat)])
```

Histogram from ETH GLIMMER
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
#used this to find out the that the two groups are 0 1 not high low, not sure why
#groups <- c("0","1")
#group_model <- mod_intuitive_eth
#Mit = as_tibble(coef(group_model, simplify=T)) %>%
#    mutate(definition = rownames(coef(group_model, simplify=T)[[groups[1]]]$items),
#           group1 = groups[1],
#          group2 = groups[2])

mm_eth <- extract_group_df(mod_intuitive_eth, groups=c("Nonwhite","White"))
eth_hist <- item_difficulty_difference_histogram(mm_eth)

orig_eth <- get_extreme_item_difficulty_differences(mm_eth, 2.25) # min:-0.97  max:1.08  (mean + 2SD) mean:0.051
eth_hist
```

```{r eth-diff-table, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
eth_items_to_remove <- orig_eth %>% filter(extreme==T)

eth_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(ses_bias = case_when(d_diff < 0 ~ "Non-White", d_diff > 0 ~ "White")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:40%;'")
```
List of Eth Items that show DIF outside of the 2SD range. (`r nrow(eth_items_to_remove)` that show DIF)

```{r, eval=F}
#do we need to have removed NA sex kids for the ses model
d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))
#might need this
d_demo <- d_demo %>% 
  filter(!is.na(eth_group)

eth_group = as.character(d_demo$eth_group)

eth_words <- c(eth_items_to_remove$definition) 
#remove words
d_mat_eth <- d_mat[, ! names(d_mat) %in% eth_words, drop = F]
# there are more IDs in d_mat_eth than in d_demo; remove those
d_mat_eth <- d_mat_eth %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_eth$data_id
d_mat_eth$data_id = NULL
d_mat_eth = data.matrix(d_mat_eth)
row.names(d_mat_eth) = sids

#refit model now with main DIF items removed
mod_intuitive_ethn <- fit_mod_intuitive(d_mat_eth, eth_group)
#recieved this note: NA values in group removed, along with associated rows in data
#may need to run the "might need this" line to remove those NA kids ahead of time
plot_glimmer(mod_intuitive_ethn, colnames(d_mat_eth)[2:ncol(d_mat_eth)])

```
```{r show-new-eth-glimmer, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
eth_words <- c(eth_items_to_remove$definition) 
#remove words
d_mat_eth <- d_mat[, ! names(d_mat) %in% eth_words, drop = F]
# there are more IDs in d_mat_eth than in d_demo; remove those
d_mat_eth <- d_mat_eth %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_eth$data_id
d_mat_eth$data_id = NULL
d_mat_eth = data.matrix(d_mat_eth)
row.names(d_mat_eth) = sids

plot_glimmer(mod_intuitive_ethn, colnames(d_mat_eth)[2:ncol(d_mat_eth)])
```

New Histogram with the DIF Items removed
```{r, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
mm_ethn <- extract_group_df(mod_intuitive_ethn, groups=c("Nonwhite","White"))
eth_histn <- item_difficulty_difference_histogram(mm_ethn)

orig_ethn <- get_extreme_item_difficulty_differences(mm_ethn, 2.25) # min:-0.97  max:1.08  (mean + 2SD) mean:0.051
eth_histn
```

```{r eth-diff-table-new, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
new_eth_items_to_remove <- orig_ethn %>% filter(extreme==T)

new_eth_items_to_remove %>% filter(extreme==T) %>%
  arrange(d_diff) %>% 
  select(-d_diff_abs, -extreme, -a1, -group1, -group2, -d_g1, -d_g2) %>%
  mutate(eth_bias = case_when(d_diff < 0 ~ "Non-White", d_diff > 0 ~ "White")) %>%
  kable(format = "html", digits = 2, table.attr = "style='width:60%;'")
```
List of new Eth Items that show DIF outside of the 2SD range. (`r nrow(new_eth_items_to_remove)` that show DIF)

```{r, race-table}
# breakdown white/non-white
```

```{r, eval=FALSE}
#save all the new models
save(mod_intuitive_sexn, mod_intuitive_ethn, mod_intuitive_sesn, mod_intuitive_sexn_n,
       file="data/glimmer_prodWS_model_noExt.Rds")

```

# Discussion

Where to go from here -- other languages, regional variation, 
- using adult word vectors to find unbiased items?