### 2022-01-24 paper/cogsci.Rmd
edits from Nathan, new refs

---
title: "Estimating demographic bias on tests of children's early vocabulary"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George Kachergis (kachergis@stanford.edu)}
    \AND {\large \bf Nathan Francis (nathan99@stanford.edu)}
    \AND {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, Stanford Unviersity \\ Stanford, CA 94305 USA
    }

abstract: >
    Children's early language skill has been linked to later educational outcomes, making
    it important to accurately measure early language. Parent-reported instruments such as the 
    Communicative Development Inventories (CDIs) have been shown to provide valid, consistent 
    measures of children's aggregate early language skill. However, CDIs contain
    hundreds of vocabulary items, some of which may not be heard (and thus learned) equally
    often by children of varying backgrounds. This study used a database of American English 
    CDIs to identify words that showed strong bias for particular demographic groups of children, on 
    dimensions of sex (male vs. female), race (white vs. non-white), and maternal education 
    (high vs. low). For each dimension, we identified dozens of strongly biased items, and showed that 
    eliminating these items reduced the expected ability difference between groups. Additionally,
    we investigated how well the relative frequency of words spoken to young girls vs. boys predicted 
    sex-based word learning bias, and discuss possible sources of demographic bias in early word learning.

keywords: >
    language acquisition; word learning; measuring instrument bias; demographics;

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(here)
library(mirt)
library(ggpubr)
library(ggrepel)

theme_set(theme_classic(base_size=10)) 
```

# Introduction

Researchers, clinicians, and parents have long been fascinated with the surprising speed and variability in the growth of young children's vocabulary.
Children's early vocabulary growth is assumed to reflect not only their exposure to child-directed speech, but also the varying difficulty of different types of words, and individual differences in the aptitude of the child -- including potential language deficits.
Children show both consistency in some skills across development, as well as significant influence from external factors.
For example, @bornstein2016stability found stability in core language skills across 10 years of children's development, despite changes in maternal income and education over the study period.
Yet maternal education, often used as a proxy for socioeconomic status (SES), has also been found to be associated with children's language processing and vocabulary by 18 months [@fernald2013ses], and to be predictive of later educational outcomes [@marchman2008speed; see @schwab2016 for a review].
Demographic factors are also predictive of language skill: first-born children tend to outpace their siblings, and female children tend to have better language skills than their age-matched male counterparts [@fenson1994] -- a sex-based verbal advantage that continues through high school [see @petersen2018gender for a review].

However, it is also difficult to get a complete measure of young children's language skill: long recordings are prohibitively difficult to collect and transcribe [see @roy2015predicting, the exception that proves the rule], and yet any short recording (e.g., a 1-hour play session) will elicit only a small proportion of the words and constructions that children know. 
Thus, researchers of early word learning have constructed tests with hundreds of words, intentionally oversampling words that are more likely to be known by young children.

We focus on the MacArthur-Bates Communicative Development Inventories [CDIs; @Fenson2007], a set of parent-reported measures of children's productive and receptive language skills, which offer a low-cost and reliable way to estimate children's early language skills [@fenson1994].
CDIs have shown good predictive validity [e.g., @fenson1994; @bornstein2012stability; @duff2015infant].
Our primary focus is the vocabulary checklist portion of the CDI Words & Sentences (CDI:WS) form, comprised of 680 early-learned words across 22 categories (e.g., animals, vehicles, action words, pronouns) selected to assess the productive vocabulary of children 16 to 30 months of age.
For each item on the CDI:WS, caregivers are asked to respond whether the target child has been heard to say (i.e. produce) the given item.[^1]
Children's total vocabulary score on the CDI:WS is tightly correlated with other facets of early language (e.g., grammatical competence and gesture), suggesting that the language system is "tightly woven" [@frank2021].
Due to these desirable properties, CDIs have been adapted to dozens of languages, and a central repository of CDI data contributed from all over the world has been created [Wordbank; @frank2017; @frank2021].

[^1]: The CDI: Words and Gestures form includes a subset of the CDI:WS items and targets children 12 to 18 months of age, measuring both comprehension and production.

Inspired by the utility and widespread use of the CDI, researchers have recently been using psychometric models on CDI data to construct short, adaptive tests to reliably assess language ability using only a small subset of the CDI items [e.g., @mayor2019; @kachergis2021cat].
These psychometric models typically come from the Item-Response Theory (IRT) framework [@Baker2001], which assumes that not only test-takers (here, children) have normally-distributed ability, but that items (words on the CDI) have normally-distributed difficulty.
The very efficacy of these IRT-based models depends on words varying in difficulty, and hence being more/less informative of the ability level of different individuals.
For example, asking whether a 22-month-old produces the word "ball" is far less informative of that child's language ability than asking whether they produce "table", as 96% of 22-month-olds can produce the former, while only 47% produce the latter.

While it is quite reasonable to expect that some CDI words are "easier" (i.e., more likely to be known by children) than others, and even to use these varying difficulties to predict variation in children's language ability, the use of psychometric models highlights the possibility that some CDI items may function differently (i.e., be more/less widely known) for different groups of children.
The idea that some items on a test may show bias, favoring one group over another, is known as Differential Item Function (DIF; @holland1993differential]. 
On any given test, it is clearly undesirable to have more items favoring one group (say, children from rural households) over another (urban children), as the test will overestimate the ability of test-takers in the former (rural) group -- and not because of any underlying mean difference in ability between the groups, but simply because the test is unfair [@camilli2013ongoing].
A variety of statistical methods for detecting DIF have been proposed, and investigations have in several instances identified DIF for many items on tests favoring one group over another (e.g., rural vs. urban).
Our goal here is to test the items on the American CDI:WS for DIF along three main axes: sex (male vs. female), maternal education (no more than secondary vs. at least some college), and race (white vs. non-white).

The outline of this paper is as follows.
First, we introduce the Wordbank data and the IRT model, and use it examine the overall size of demographic differences in early word learning.
We then fit the IRT model to each group along each demographic dimension, and examine the item parameters for evidence of DIF, noting in particular how many items are significantly biased in favor of each group.
We identify a set of suspect items to eliminate (or in the future, replace), and provide updated estimates of the effect size of these demographic variables, were the biased items to be eliminated.
<!--Next, using a corpus of child-directed speech, we examine the extent to which variation in word frequencies heard by boys vs. girls predicts sex-based DIF.-->
Finally, we provide recommendations for next steps to be taken to identify and replace biased items on the CDI.

# Methods

## Vocabulary Data

```{r import-data}
load(here("data/en_ws_production.Rdata")) # d_demo, d_mat, en_wg (items)

#Add this exclusion to the Analysis 
removed <- d_demo %>%
  filter(comprehension!=0)
#this is the actually d_demo we used so we must use this in our report
d_demo <- d_demo %>% 
  #filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))
```

### Participants

```{r, echo=F}
sex_tab <- table(d_demo$sex)
ses_tab <- table(d_demo$ses_group)
eth_tab <- table(d_demo$ethnicity)
eth2_tab <- table(d_demo$eth_group)
momed_tab <- table(d_demo$mom_ed)
```

We analyze parent-reported Wordbank data from `r nrow(d_demo)` American English CDI: Words & Sentences administrations for children `r min(d_demo$age)` to `r max(d_demo$age)` months of age [@frank2017; @frank2021].
Full demographic data are not reported in some datasets contributed to Wordbank: sex was available for `r nrow(d_demo)-sum(is.na(d_demo$sex))` children, race/ethnicity was available for `r nrow(d_demo)-sum(is.na(d_demo$ethnicity))` children, and maternal education (a proxy for socioeconomic status; SES) was available for `r nrow(d_demo)-sum(is.na(d_demo$ses_group))` children.

The analysis of sex-based differences included CDI administrations from `r sex_tab["Female"]` female and `r sex_tab["Male"]` male children.
The analysis of race-based differences included data from `r eth_tab["White"]` white, `r eth_tab["Asian"]` Asian, `r eth_tab["Black"]` Black, `r eth_tab["Hispanic"]` Hispanic, and `r eth_tab["Other"]` "Other" children. 
Due to sparse data for many categories, we binarized participants' race/ethnicity as White (`r eth2_tab["White"]`) or Non-white (`r eth2_tab["Nonwhite"]`), recognizing that there may be important variation between groups that this will fail to capture. 
Data for the maternal education analysis included CDI data from children whose mother's had the following levels of education: `r momed_tab["Primary"]` with no more than primary school education, `r momed_tab["Some Secondary"]` with some secondary school, `r momed_tab["Secondary"]` with no more than secondary school, `r momed_tab["Some College"]` with some college, `r momed_tab["College"]` with no more than a college degree, `r momed_tab["Some Graduate"]` with no more than some graduate school, and `r momed_tab["Graduate"]` with a graduate degree.
Again due to data sparsity, we binarized the `r ses_tab["high"]` children whose mothers had at least some college or more as high maternal education (high-ME), and those whose mothers had at most high school (`r ses_tab["low"]` children) as low maternal education (low-ME).

## Rasch Model

The Rasch model, also known as the 1-parameter logistic (1PL) model, is the simplest Item Response Theory model, and is thus the easiest to use to investigate potential differences in item function across different groups of participants.
The goal of the Rasch model is to jointly estimate for each child $j$ a latent ability $\theta_j$, and for each item $i$ a difficulty parameter $b_i$.
In the model, the probability of child $j$ knowing (i.e., producing or understanding) a given item $i$ is

$$P_{i}(x_i = 1 | b_{i},\theta_j ) = \frac{1}{1 + e^{-D(\theta_j - b_i )}}$$

where $D$ is a constant scaling parameter ($D=1.702$) which makes the logistic closely match the ogive function in traditional factor analysis [@R-mirt; @reckase2009].
Child ability ($\theta$) and item difficulty ($b$) distributions are standardized (i.e., mean of 0), and expected to be normally-distributed. 
Children with high latent ability ($\theta$) will be more likely to produce any given item than children with lower latent ability, and more difficult items will be produced by fewer children (at any given $\theta$) than easier items.

In the multigroup Rasch model, an item's difficulty is allowed to vary depending on which group an individual child belongs to. 
For example, in the sex-based multigroup model, item $i$'s difficulty is $b_{i}^{female}$ for females, and $b_{i}^{male}$ for males.
In the quest to identify DIF, a multigroup Rasch model will be fitted for each demographic dimension of interest (sex, maternal education, and ethnicity), and we will examine the between-group difficulty difference for each item (e.g., $d_i = b_{i}^{female} - b_{i}^{male}$).
If there is no DIF for a given item, then $d_i \approx 0$ as the two groups find the item equally difficult.

# Results

We will first examine the size of demographic effects on language ability in a baseline Rasch model fitted without regard to demographic group.
We then fit a multigroup Rasch model for each examine evidence for DIF, assuming that 
Finally, we prune the CDI of items showing varying degrees of DIF

```{r colors}
# color-blind palette
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
sex_colors = cbp1[2:3]
eth_colors = cbp1[4:5]
ses_colors = cbp1[6:7]
```

```{r load-data}
load(here("data/glimmer_prodWS_models.Rds"))
load(here("data/eng_Ws_prod_mod_1pl.Rds")) # m0h - baseline Rasch model, with coefs_1pl and fscores_1pl
#cor.test(d_demo$production, d_demo$ability) # .972
d_demo <- d_demo %>% left_join(fscores_1pl)

#d_demo %>% filter(!is.na(sex)) %>%
#  ggplot(aes(x=sex, y=ability)) +
#  geom_boxplot() + theme_classic()

sex_m <- d_demo %>% filter(!is.na(sex), !is.na(ability), age<26, age>22) %>% 
  group_by(sex) %>% summarise(ability=mean(ability),
                              production=mean(production),
                              n=n()) # 394 female, 408 male

ses_m <- d_demo %>% filter(!is.na(ses_group), !is.na(ability), age<26, age>22) %>% 
  group_by(ses_group) %>% summarise(ability=mean(ability),
                                    production=mean(production),
                                    n=n()) # 1014 high-SES, 140 low-SES

eth_m <- d_demo %>% filter(!is.na(eth_group), age<26, age>22) %>% 
  group_by(eth_group) %>% summarise(ability=mean(ability),
                                    production=mean(production),
                                    n=n()) # 144 nonwhite, 603 white

# sex_m[1,2:3] - sex_m[2,2:3] # average female vs. male advantage at 23-25 months of age: 0.78 theta, 69 words
# ses_m[1,2:3] - ses_m[2,2:3] # high vs. low advantage: 0.33 theta, 33 words
# eth_m[2,2:3] - eth_m[1,2:3] # white vs. nonwhite advantage: 0.52 theta, 45

# regression - report coefficients again after we cut some items?
d_demo$eth_group = factor(d_demo$eth_group, levels=c("Nonwhite","White"))
d_demo$sex = factor(d_demo$sex, levels=c("Male","Female"))
d_demo$ses_group = factor(d_demo$ses_group, levels=c("low","high"))
d_demo <- d_demo %>% mutate(age_sc = scale(age, scale=F))
eth_reg <- summary(lm(ability ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 white-, age*white+
ses_reg <- summary(lm(ability ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg <- summary(lm(ability ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

#ab_reg <- summary(lm(ability ~ age_sc * (eth_group + ses_group + sex), data=d_demo)) # R^2=.49
#prod_reg <- summary(lm(production ~ age_sc * (eth_group + ses_group + sex), data=d_demo)) # R^2=.51
```

## Demographic effects in baseline Rasch model

To contextualize the later DIF results, we first fitted a baseline Rasch model to the entire dataset, without demographic information.
Figure 1 shows children's language ability vs. age by demographic group from the baseline Rasch model, which assumes no DIF (i.e., equal item parameters for all groups).
A linear regression for each demographic group, with age (centered) and its interaction, showed significant effects in the expected directions, as can also be seen in the figure. 
Female children had higher language ability than male children ($\beta=0.56$, $p<.001$), with no significant interaction with age ($\beta=0.02$, $p=.10$).
High-SES children had higher language ability than low-SES children ($\beta=0.23$, $p=.02$), an advantage that grew with age ($\beta=0.11$, $p<.001$).
White children had higher language ability than non-white children ($\beta=0.50$, $p<.001$), an advantage that grew with age ($\beta=0.08$, $p<.001$).
We will re-examine these demographic regressions after trimming items showing extreme DIF.

<!-- Although the demographic language ability advantages go in the expected directions (i.e., favoring female, high-SES, and white children), note that magnitude of the differences vary somewhat across the age range, especially for SES and ethnicity.
Thus, although the IRT models are fitted to the entire age range, our comparisons of average ability will focus on the middle of the CDI:WS' intended age range, where there is great variation in the size and composition of children's vocabulary.
At ages 23-25 months, the average female vs. male advantage is an ability difference of $\theta_{f-m} = 0.78$, amounting to females knowing 69 more words, on average.
The average advantage for high-SES children at 23-25 months is $\theta_{high-low} = 0.33$, amounting to 33 more produced CDI items.
The average advantage for white vs. non-white children in this age range is $\theta_{w-nonw} = 0.52$, an average of 45 more produced CDI items.
-->

```{r baseline-ability-vs-age, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.3, fig.height=6, set.cap.width=T, num.cols.cap=1, fig.cap = "Language ability vs. age by demographic group, from the Rasch model."}
p1 <- d_demo %>% filter(!is.na(sex)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=sex)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") + 
  scale_color_discrete(type=sex_colors)
  #theme(legend.position = "top")

p2 <- d_demo %>% filter(!is.na(ses_group)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=ses_group)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") +
  scale_color_discrete(name = "SES", type=ses_colors) #+ 
  #theme(legend.position = "top")

p3 <- d_demo %>% filter(!is.na(eth_group)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=eth_group)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") + 
  scale_color_discrete(name = "ethnicity", type=eth_colors) #+ 
  #theme(legend.position = "top")

ggarrange(p1, p2, p3, ncol=1)
```

## Identifying biased CDI items

The first step we take to identify CDI items with DIF is to create GLIMMERs (Graphs of Logits Imputed Multiply with Means Equal; @stenhaug2021treading), which visualize between-item variation in group performance differences.
These parameters are drawn from a fitted multigroup Rasch model for each demographic variable (sex, SES, and race), with the assumption that the mean language ability in each group is the same (e.g. for sex, $\mu_{male}=\mu_{female}=0$), thus pushing all between-group variation into the item difficulty parameters.
For the case of sex, where we believe $\mu_{female}>\mu_{male}$, this means we may expect to find many items with difficulty $b_i^{female} < b_i^{male}$, but we can still examine the distribution of differences in item difficulty ($d_j = b_i^{male} - b_i^{female}$) for outliers.
To give analysts a sense of the uncertainty about the existence of DIF, GLIMMER plots show distributions of parameter differences rather than point estimates. 
These distributions are generated by drawing 10,000 imputations from the item parameter covariance matrix.

Figures 2-4 show GLIMMERs for a selection of CDI items for sex, SES, and race, respectively.
The full GLIMMERs, with all 680 CDI:WS items, are available on [OSF](https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a), but were too large to include here.
Nonetheless, it is important to inspect the full plots, for if there is a cluster of items in a GLIMMER, the analyst may conclude that these items are strong candidates for DIF on that dimension.
For example, there is some clustering at the top and bottom of Figure 2: at the top, "vagina", "tights", "dress (object)", and "doll" form a cluster of items that are much easier for females, while at the bottom, "penis" stands out as much easier for males.
For the maternal education and race GLIMMERs (Figures 3 and 4)--and in the rest of the full sex GLIMMER, there are not clusters, but rather a continuum of smoothly varying differences with overlap. 
This makes identifying items with DIF quite difficult, as different methods are likely to yield inconsistent results [@stenhaug2021treading].
Nonetheless, we describe the distributions of item-level group difficulty differences, and identify extreme values, flagging items with difficulty differences that were outside 2 standard deviations of the mean of that distribution.

```{r get-dif-dfs}
extract_group_df <- function(group_model, groups=c("Male","Female")) {
  Mit = as_tibble(coef(group_model, simplify=T)[[groups[1]]]$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)[[groups[1]]]$items),
           group1 = groups[1],
           group2 = groups[2]) %>%
    select(-g, -u) %>% 
    rename(d_g1 = d)
  Fit = as_tibble(coef(group_model, simplify=T)[[groups[2]]]$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)[[groups[2]]]$items)) %>%
    select(-g, -u) %>%
    rename(d_g2 = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_g2 - d_g1,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

get_extreme_item_difficulty_differences <- function(mm, sd_thresh=2, verbose=F) {
  #yfit <- dnorm(mm$d_diff, mean = mean(mm$d_diff), sd = sd(mm$d_diff)) 
  mm <- mm %>% mutate(d_diff = d_g2 - d_g1)
  max_dif = mean(mm$d_diff) + sd_thresh*sd(mm$d_diff)
  min_dif = mean(mm$d_diff) - sd_thresh*sd(mm$d_diff)
  mm <- mm %>% mutate(extreme = ifelse((d_diff > max_dif) | (d_diff < min_dif), T, F))
  if(verbose) print(paste("mininum difference:",round(min_dif,3), "maximum difference:",round(max_dif,3)))
  return(mm)
}

mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
mm_ses <- extract_group_df(mod_intuitive_ses, groups=c("low","high"))
mm_eth <- extract_group_df(mod_intuitive_eth, groups=c("Nonwhite","White"))
```

```{r train-pruned-models, eval=F}
# varying thresholds
prune_dat <- tibble()
for(thresh in seq(.25, 3.0, .25)) {
  tmp_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=thresh) # min: -.53 max: 1.06
  tmp_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=thresh) # min: -1.07 max: 1.13
  tmp_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=thresh) # min: -.64 max: 1.56
  prune_defs <- unique(c(subset(tmp_ses, extreme)$definition, 
                       subset(tmp_eth, extreme)$definition,
                       subset(tmp_sex, extreme)$definition))

  # prune the columns
  keep_cols <- which(!is.element(names(d_mat), prune_defs))
  #names(d_mat)[which(is.element(names(d_mat), prune_defs))]
  d_mat_pruned <- d_mat[,keep_cols]

  m0 <- mirt(d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(-data_id), 1, "Rasch", 
           technical=list(NCYCLES=1000))

  #coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  #  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
  fscores_1pl <- tibble(data_id = subset(d_mat_pruned, rowSums(d_mat_pruned)!=0)$data_id, 
                               ability_pr = fscores(m0, method = "MAP")[,1])

  d_demo_tmp <- d_demo %>% left_join(fscores_1pl) %>%
    mutate(age_sc = scale(age, center=T, scale=F))
  # join with d_demo, do the demo regressions again..

  eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo_tmp))$coefficients # R^2=.48 
  ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo_tmp))$coefficients # R^2=.56
  sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo_tmp))$coefficients # R^2=.59

  prune_dat = bind_rows(prune_dat, 
                        tibble(SDthreshold=thresh, 
                               extreme_sex=nrow(subset(tmp_sex, extreme)),
                               extreme_ses=nrow(subset(tmp_ses, extreme)),
                               extreme_eth=nrow(subset(tmp_eth, extreme)),
                               extreme_total = length(prune_defs),
                               sex_beta = sex_reg_pr[3,1],
                               ses_beta = ses_reg_pr[3,1],
                               eth_beta = eth_reg_pr[3,1]))
}

save(prune_dat, file=here("data/pruned_demo_coefs.Rdata"))
```

```{r}
# 2SD threshold
mm_sex <- get_extreme_item_difficulty_differences(mm_sex) # min: -.53 max: 1.06
mm_ses <- get_extreme_item_difficulty_differences(mm_ses) # min: -1.07 max: 1.13
mm_eth <- get_extreme_item_difficulty_differences(mm_eth) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex, extreme),
                     subset(mm_ses, extreme),
                     subset(mm_eth, extreme)) 
# length(unique(all_extreme$definition)) 
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

# overlap between extreme SES and race items:
#intersect(subset(mm_ses, extreme)$definition, subset(mm_eth, extreme)$definition)
```

```{r pruned-coefs, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.4, fig.height=4.75, set.cap.width=T, num.cols.cap=1, fig.cap = "Size of demographic effects (regression coefficients) vs. the threshold of pruned items."}
load(here("data/pruned_demo_coefs.Rdata"))

# prune_dat %>% rename(c(beta_sex=sex_beta, beta_ses=ses_beta, beta_eth=eth_beta)) %>%
#   pivot_longer(c("extreme_sex","extreme_ses","extreme_eth","beta_sex","beta_ses","beta_eth"), 
#                            names_to=c("type", "demo"), names_sep="_") %>%
#   pivot_wider(names_from=type, values_from=value) %>%
#   mutate(Dimension = ifelse(demo=="ses", "maternal ed", ifelse(demo=="eth", "race", demo))) %>%
#   ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + facet_wrap(. ~ Dimension, scales="free") + 
#   geom_point() + theme(legend.position = "bottom") + # geom_line() +
#   labs(color="Excluded items") + 
#   xlab("Exclusion Threshold (SDs)") + ylab("Regression coefficient (beta)")

pdat <- prune_dat %>% rename(c(beta_sex=sex_beta, beta_ses=ses_beta, beta_eth=eth_beta)) %>%
  pivot_longer(c("extreme_sex","extreme_ses","extreme_eth","beta_sex","beta_ses","beta_eth"), 
                           names_to=c("type", "demo"), names_sep="_") %>%
  pivot_wider(names_from=type, values_from=value) %>%
  mutate(Dimension = ifelse(demo=="ses", "maternal ed", ifelse(demo=="eth", "race", demo)))

pr1 <- pdat %>% filter(Dimension=="sex") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + theme(legend.position = "bottom") + 
  labs(color="Excluded items") + ylim(-.7, -.4) + ggtitle("Sex") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

pr2 <- pdat %>% filter(Dimension=="maternal ed") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + theme(legend.position = "bottom") + 
  labs(color="Excluded items") + ylim(-.3, -.1) + ggtitle("Maternal Education") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

pr3 <- pdat %>% filter(Dimension=="race") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + theme(legend.position = "bottom") + 
  labs(color="Excluded items") + ylim(.35, .55) + ggtitle("Race") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

ggarrange(pr1, pr2, pr3, ncol=1, common.legend = T)
```

### Sex
For sex, the median difficulty difference (male-female) was `r round(median(mm_sex$d_diff),2)` (M=`r round(mean(mm_sex$d_diff),2)`, sd=`r round(sd(mm_sex$d_diff),2)`), with `r length(which(mm_sex$d_diff>0))`/680 items being easier for females than males.
The fact that the bulk of this distribution favors females seems to confirm that the female language advantage is pervasive across the CDI:WS, and is thus likely to be a real advantage.

### Maternal Education
For maternal education (ME), the median difficulty difference (low-high) was `r round(median(mm_ses$d_diff),2)` (M=`r round(mean(mm_ses$d_diff),2)`, sd=`r round(sd(mm_ses$d_diff),2)`), with `r length(which(mm_ses$d_diff>0))`/680 items being easier for high-ME than low-ME children. 
With the mean and median difficulty differences close to 0, and roughly half of the words favoring each ME group, it is tempting to conclude that the CDI:WS items are somewhat balanced with respect to ME.

### Race
For race, the median difficulty difference (nonwhite-white) was `r round(median(mm_eth$d_diff),2)` (M=`r round(mean(mm_eth$d_diff),2)`, sd=`r round(sd(mm_eth$d_diff),2)`), with `r length(which(mm_eth$d_diff>0))`/680 items being easier for white than non-white children, revealing a fairly pervasive advantage for white children on CDI items.

This identified 25 items with extreme sex-based difficulty differences, only 7 of which were easier for females, with the other 18 items favoring males. <!-- thus, removing these extrema would actually lower male children's vocabulary scores..] -->
For SES, 39 extrema were identified, only 13 of which were easier for low-SES children.
For race, 31 extrema were identified, only 11 of which were easier for non-white children.
Thus, while if anything the bulk of the extrema in the sex-based model advantaged the lower language ability group (males), the extrema in the other two models mostly favored the advantaged groups: 26/39 extreme items in the SES model favored high-SES children, and 20/31 items in the race model favored white children.
17 CDI items were identified as extrema in more than one model: "tractor" and "vroom" were extreme in all three models, "choo choo" was extreme in both SES and sex models, "give me five!" was extreme in both race and sex models, and 13 other items were extreme in both SES and race models ("grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy", "pet's name", "up", and "so").
In our final analysis, we explore the effect of pruning the 15 items that were extreme in both SES and race models.

```{r items-to-remove}
prune_defs <- c("tractor", "vroom", "grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy*", "pet's name", "up", "so") # also "choo choo" and "give me five!" ?
prune_cols <- which(!is.element(names(d_mat), prune_defs))
#names(d_mat)[which(is.element(names(d_mat), prune_defs))]
d_mat_pruned <- d_mat[,prune_cols]
```

```{r fit-pruned-model, eval=F}
m0 <- mirt(d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(-data_id), 1, "Rasch", 
           technical=list(NCYCLES=1000))

coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
fscores_1pl <- tibble(data_id = d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(data_id), 
                             ability_pr = fscores(m0, method = "MAP")[,1])
fscores_1pl$data_id = fscores_1pl$data_id$data_id
save(file = here("data/eng_Ws_prod_mod_1pl_pruned.Rds"), "m0", "coefs_1pl", "fscores_1pl")
```

```{r sex-glimmer, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.2, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the sex bias model. Words at the top are easier to learn for females, while those at the bottom are easier for males."}
#img <- png::readPNG("figs/lab_logo_stanford.png")
#grid::grid.raster(img)
knitr::include_graphics("figs/smGLIMMER_sex_prodWS.pdf")
```

```{r ses-glimmer, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the SES bias model. Words at the top are easier to learn for children from low-SES families, while those at the bottom are easier for those from high-SES families."}
knitr::include_graphics("figs/smGLIMMER_ses_prodWS.pdf")
```

```{r eth-glimmer, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the ethnicity bias model. Words at the top are easier to learn for children from low-SES families, while those at the bottom are easier for those from high-SES families."}
knitr::include_graphics("figs/smGLIMMER_eth_prodWS.pdf")
```

```{r, eval=F, include=F}
View(subset(mm_sex, extreme)) # 25 extreme (outside 2sd) - 7 favoring girls, 18 favoring boys
# possibly underestimating the female advantage?
View(subset(mm_ses, extreme)) # 39 extreme - 13 favoring low-SES, 26 favoring high-SES
View(subset(mm_eth, extreme)) # 31 extreme - 11 favoring nonwhite, 20 favoring white
# overestimating both high-SES and white advantages?
```

## Demographic effects after pruning

```{r load-pruned}
# (2sd threshold)
load(here("data/eng_Ws_prod_mod_1pl_pruned.Rds"))

d_demo <- d_demo %>% left_join(fscores_1pl)
# join with d_demo, do the demo regressions again..

eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 
ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

#white=.48
#highSES=0.20
#female=0.57

# table proportion of DIF items by lexical class -
# if due to environmental 
```

We identified 76 CDI items that function quite differently across demographic groups, and chose to prune just the 15 items with large difficulty differences for both SES and race. 
After pruning just these 15 items from the CDI:WS and re-fitting a Rasch IRT model, regressions predicting ability with each demographic variable (and age, centered) showed reduced main effects of SES ($\beta=0.20$ vs. $\beta=0.23$) and race ($\beta=0.48$ vs. $\beta=0.50$).
Although these are small reductions in the SES and race effect sizes, we find them striking as we pruned only 2.2% of the CDI:WS items (and only 20% of the 76 extrema we identified), and did not specifically prune items that were biased against the disadvantaged groups (non-white, low-SES).

```{r demo-ability, include=T, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=7.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Scatterplots showing item difficulty per demographic group along each dimension."}

p1 <- mm_sex %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) + # color=extreme
  xlab("Male Difficulty") + ylab("Female Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

p2 <- mm_ses %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) +
  xlab("Low-ME Difficulty") + ylab("High-ME Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

p3 <- mm_eth %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) +
  xlab("Non-white Difficulty") + ylab("White Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

# p1 <- mm_sex %>% 
#   ggplot(aes(x=d_diff)) + # , color=extreme, fill=extreme
#     geom_density(alpha=.2) + xlab("Ability") 
# 
# p2 <- mm_ses %>%
#   ggplot(aes(x=d_diff)) + 
#     geom_density(alpha=.2) + xlab("Ability") 
# 
# p3 <- mm_eth %>%
#   ggplot(aes(x=d_diff)) + 
#     geom_density(alpha=.2) + xlab("Ability")

ggarrange(p1, p2, p3, ncol=1)
```

## Relating child-directed speech to demographic bias

Demographic differences in language ability are likely to be at least partially explained by differences in linguistic input received by children in different groups.
Indeed, input quantity (total daily tokens) and some measures of quality (e.g., lexical diversity: ratio of word types vs. tokens) have been predictive of language learning outcomes in some demographic studies [@rowe2009differences; @huttenlocher1991early].
Here, we investigated the extent to which word frequency in child-directed speech to male vs. female children was predictive of the amount of DIF shown by CDI items. 
Similar to the approach taken by @braginsky2016gender, we used the CHILDES corpus of transcripts from dyadic play sessions [@macwhinney2000childes], which are labeled with the sex of the target child, but not other demographic variables.
(Number of males, number of females, total word counts, a couple examples...)

Overall, the correlation between the proportion of times a word was spoken to a female child and the size of the female (vs. male) advantage for that CDI word was modest, but significant ($r=0.19$, $p<.001$).

```{r, include=T, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "CHILDES frequency of input to male vs. female children."}
type_freqs = readRDS("../data/type_freqs.rds")
type_counts = readRDS("../data/type_counts.rds")
type_counts_smoothed = readRDS("../data/type_counts_smoothed.rds")

type_counts_smoothed <- type_counts_smoothed %>%
  pivot_wider(id_cols=uni_lemma, names_from=sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male),
         FtoM = female / male)

# how biased are the extreme items we suggest removing in the corpus counts?
#sex_items_to_remove$definition

tmp <- left_join(mm_sex, type_counts_smoothed, by=c("definition"="uni_lemma"))

tmp %>% #filter(extreme==T) %>%
  ggplot(aes(x=propFemale, y=d_diff, color=extreme)) + 
  geom_point(alpha=.5) + theme_bw() +
  ylab("Female - Male Word Difficulty") +
  xlab("Proportion of Utterances to Females in CHILDES") +
  theme(legend.position = "bottom")

#subset(tmp, extreme==T & propFemale > .47 & propFemale < .53) # hose & dump

#cor.test(tmp$propFemale, tmp$d_diff) # .19, p<.001

```

# Discussion

We investigated the CDI:WS, a popular parent-report measure of children's early vocabulary, for potential demographic bias, examining the distribution of words' estimated difficulties for high- vs. low-SES children, females vs. males, and for white vs. non-white children.
The IRT-based analysis revealed differential item functioning (DIF) for many items along each demographic dimension, but only in the case of sex were clear clusters of items that were more well-known to females (including feminine clothing and genitalia), and a clear item that was more well-known to males (male genitalia). 
For the rest of the items, and for SES- and race-based analysis, there was a smooth continuum of DIF, making the boundary of true DIF subjective, as this would rely on knowing the true difference in language ability between groups--which we reciprocally estimate from instruments like the CDI.
To move forward, we identified candidate DIF items by looking at the extremes of each distribution, and found that for SES and race, the majority of these extrema (66% in both cases) were easier for the majority demographic group (i.e., high-SES, white), meaning that if these items were removed, SES and race effects on language ability as measured by the CDI:WS would decrease.
In contrast, the majority of the extrema in the sex-based DIF analysis were easier for the disadvantaged group (males): if these extrema are removed, the female language advantage on the CDI:WS will likely increase.
To confirm these implications, we pruned a very small number of CDI items (15 of 680) that were DIF extrema in both SES- and race-based analyses, and found that the effect of these dimensions on language ability was reduced.

## The difficulty of DIF

However, DIF is fundamentally difficult to identify for multiple reasons [for an overview, see @stenhaug2021treading].
First, most techniques to identify DIF rely on defining a set of "anchor" test items that are assumed to be equally difficult (i.e., unbiased) for both groups of interest.
Identifying anchor items is at best fraught when there may in fact be a difference in ability between groups (e.g., the female language advantage), and is further confounded when the magnitude of this ability difference is unknown.
A reason that DIF is particularly tricky in measuring children's early language ability is that there is a finite universe of early-learned words to choose from--and we may expect many of them to be biased for various environmental reasons (e.g., children in Florida may not use mittens or skis).
Hence, the presence of DIF on a wide variety of items may not indicate the presence of bias; it could indicate that one demographic has a higher average ability level than the other.
For example, let male language ability be drawn from a standard normal $\mu_{male} ~ N(0,1)$, with female language ability slightly higher, on average ($\mu_{female} = \mu_{male} + 0.1$).
Then we would expect items to be an average of $0.1$ easier for females than for males, and we might identify items that are instead easier for males than females as showing undesirable DIF.
And yet, without knowing the actual ability difference between two demographic groups--for which we also rely upon our tests, it is difficult to adjudicate which items show DIF, and which do not.

This investigation is only a first step in measuring demographic bias in the items on the CDI:WS.
In aggregate, it seems to suggest that we may be slightly overestimating the magnitude of differences in language ability along the dimensions of SES and race, and perhaps underestimating the female advantage.
However, the best way to estimate endogenous differences in ability (rather than exogenous differences, e.g. in the language environment) would be to conduct controlled, in-lab experiments measuring children's ability to learn novel words.
Future research may also look into the specific items that are easier for one group than another, and determine whether these differences can be accounted for by different environmental contexts. 
For example, it is striking that many of the extrema favoring high-SES children are animals and animal sounds (e.g., "grrr", "quack quack", "woof woof", "baa baa", "duck", "sheep", "giraffe", "zebra"): do high-SES households visit the zoo more, or do they often engage in other activities related to naming animals and noises they make?
If certain high-SES activities are driving early word learning for these children, what are the activities (and associated vocabulary) that low-SES households are instead engaging in?
A truly fair test of children's early vocabulary would contain a representative sample of words from all activities that children engage in, across demographic groups.

# Acknowledgements

[Redacted for anonymous review.]
<!--We thank members of the Language and Cognition lab for their feedback.-->

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent