### 2022-01-31 paper/cogsci.Rmd
discussion rewrite, pruned items Venn diagram

---
title: "Estimating demographic bias on tests of children's early vocabulary"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf George Kachergis (kachergis@stanford.edu)}
    \AND {\large \bf Nathan Francis (nathan99@stanford.edu)}
    \AND {\large \bf Michael C. Frank (mcfrank@stanford.edu)} \\ Department of Psychology, Stanford Unviersity \\ Stanford, CA 94305 USA
    }

abstract: >
    Children's early language skill has been linked to later educational outcomes, making
    it important to accurately measure early language. Parent-reported instruments such as the 
    Communicative Development Inventories (CDIs) have been shown to provide valid, consistent 
    measures of children's aggregate early language skill. However, CDIs contain
    hundreds of vocabulary items, some of which may not be heard (and thus learned) equally
    often by children of varying backgrounds. This study used a database of American English 
    CDIs to identify words that showed strong bias for particular demographic groups of children, on 
    dimensions of sex (male vs. female), race (white vs. non-white), and maternal education 
    (high vs. low). For each dimension, we identified dozens of strongly biased items, and showed that 
    eliminating these items reduced the expected ability difference between groups. Additionally,
    we investigated how well the relative frequency of words spoken to young girls vs. boys predicted 
    sex-based word learning bias, and discuss possible sources of demographic bias in early word learning.

keywords: >
    language acquisition; word learning; measuring instrument bias; demographics;

output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(here)
library(mirt)
library(ggpubr)
library(ggrepel)

theme_set(theme_classic(base_size=10))

# color-blind palette
cbp1 <- c("#999999", "#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
sex_colors = cbp1[2:3]
eth_colors = cbp1[4:5]
ses_colors = cbp1[6:7]
```

# Introduction

Researchers, clinicians, and parents have long been fascinated with the surprising speed and variability in the growth of young children's vocabulary.
Children's early vocabulary growth is assumed to reflect not only their exposure to child-directed speech, but also the varying difficulty of different types of words, and individual differences in the aptitude of the child -- including potential language deficits.
Children show both consistency in some skills across development, as well as significant influence from external factors.
For example, @bornstein2016stability found stability in core language skills across 10 years of children's development, despite changes in maternal income and education over the study period.
Yet maternal education, often used as a proxy for socioeconomic status (SES), has also been found to be associated with children's language processing and vocabulary by 18 months [@fernald2013ses], and to be predictive of later educational outcomes [@marchman2008speed; see @schwab2016 for a review].
Other demographic factors are also predictive of language skill: first-born children tend to outpace their siblings, and female children tend to have better language skills than their age-matched male counterparts [@frank2021; @eriksson2012differences] -- a sex-based verbal advantage that continues through high school [see @petersen2018gender for a review].

However, it is also difficult to get a complete measure of young children's language skill: long recordings are prohibitively difficult to collect and transcribe [see the singular exception: @roy2015predicting], and yet any short recording (e.g., a 1-hour play session) will elicit only a small proportion of the words and constructions that children know. 
Thus, researchers of early word learning have constructed tests with hundreds of words, intentionally oversampling words that are more likely to be known by young children.

We focus on the MacArthur-Bates Communicative Development Inventories [CDIs; @fenson1994; @Fenson2007], a set of parent-reported measures of children's productive and receptive language skills, which offer a low-cost and reliable way to estimate children's early language skills [@fenson1994].
CDIs have shown good predictive validity [e.g., @fenson1994; @bornstein2012stability; @duff2015infant].
Our primary focus is the vocabulary checklist portion of the CDI Words & Sentences (CDI:WS) form, comprised of 680 early-learned words across 22 categories (e.g., animals, vehicles, action words, pronouns) selected to assess the productive vocabulary of children 16 to 30 months of age.
For each item on the CDI:WS, caregivers are asked to respond whether the target child has been heard to say (i.e. produce) the given item.
Children's total vocabulary score on the CDI:WS is tightly correlated with other facets of early language (e.g., grammatical competence and gesture), suggesting that the language system is "tightly woven" [@frank2021].
Due to these desirable properties, CDIs have been adapted to dozens of languages, and a central repository of CDI data contributed from all over the world has been created [Wordbank; @frank2017; @frank2021].

<!-- [^1]: The CDI: Words and Gestures form includes a subset of the CDI:WS items and targets children 12 to 18 months of age, measuring both comprehension and production. -->

<!--Inspired by the utility and widespread use of the CDI, researchers have recently been using psychometric models on CDI data to construct short, adaptive tests to reliably assess language ability using only a small subset of the CDI items [e.g., @mayor2019; @kachergis2021cat].
These psychometric models typically come from the Item-Response Theory (IRT) framework [@Baker2001], which assumes that not only test-takers (here, children being reported on) have normally-distributed ability, but that items (words on the CDI) have normally-distributed difficulty.
The efficacy of these IRT-based models depends on words varying in difficulty, in order for the test to be more informative of the ability level of each individual.
For example, asking whether a 22-month-old produces the word "ball" is far less informative of that child's language ability than asking whether they produce "table", as 96% of 22-month-olds can produce the former, while only 47% produce the latter. -->

<!--While it is quite reasonable to expect that some CDI words are "easier" (i.e., more likely to be known by children) than others, and even to use these varying difficulties to predict variation in children's language ability, the use of psychometric models highlights the possibility that some CDI items may function differently (i.e., be more/less widely known) for different groups of children. -->

While many languages show similar demographic effects on early word learning [e.g., the female advantage: @eriksson2012differences; and the first-born advantage: @frank2021], the size of these demographic effects vary across languages [see @frank2021, Ch. 6]. 
Why might we see such variation?
While it is possible that there is an underlying difference in mean language ability between demographic groups, and that cross-linguistic variation is driven by differences in the language environments of different cultures, could it also be possible that the tests are biased to some degree--and perhaps to a different degree in different languages?
The idea that some items on a test may show bias, favoring one group over another, is known in psychometrics as Differential Item Function (DIF; @holland1993differential]. 
The validity of any test with many items favoring one group over another (say, children from rural vs. urban households) is questionable, as the test may be overestimating the ability of test-takers in the former (rural) group, and/or underestimating the ability of the latter group -- and not because of any underlying mean difference in ability between the groups, but simply because the test is unfair [@camilli2013ongoing].
For example, if a vocabulary test is composed almost entirely of the names of farm equipment, the test will underestimate the knowledge of urban children, who have experience of other contexts, despite being less familiar with farm equipment than their rural peers.
Of course, there may in fact be some meaningful ability difference between demographic groups, as is likely the case for the female language advntage, which has been documented in many languages [@eriksson2012differences; @frank2021].
However, these demographic differences in early language learning have not been investigated as potentially due to DIF.
That is the contribution of the present study.
Specifically, our goal is to test the 680 vocabulary items on the American CDI:WS for DIF along three dichotomized demographic dimensions: sex (male vs. female), maternal education (no more than secondary vs. at least some college), and race (white vs. non-white).

<!-- A variety of statistical methods for detecting DIF have been proposed, and investigations have in several instances identified DIF for many items on tests favoring one group over another. -->

The outline of this paper is as follows.
First, we introduce the Wordbank data and the IRT model, and examine the overall size of demographic differences in early word learning based on the full CDI:WS.
We then fit the IRT model to both groups for each demographic factor, and look at the item parameters for evidence of DIF, noting in particular how many items are significantly biased in favor of each group, and describing them qualitatively.
Next, we use these item parameters to systematically prune biased items from the CDI:WS, and examine how the estimated size of demographic differences change as we prune more items.
<!-- We identify a set of suspect items to eliminate (or in the future, replace), and provide updated estimates of the effect size of these demographic variables, were the biased items to be eliminated.-->
Finally, using a corpus of child-directed speech, we examine the extent to which variation in word frequencies heard by boys vs. girls predicts sex-based DIF.
Based on these analyses, we provide recommendations for next steps to be taken to identify and potentially replace items on the CDI showing demographic bias.

# Methods

## Vocabulary Data

```{r import-data}
load(here("data/en_ws_production.Rdata")) # d_demo, d_mat, en_wg (items)

#Add this exclusion to the Analysis 
removed <- d_demo %>%
  filter(comprehension!=0)
#this is the actually d_demo we used so we must use this in our report
d_demo <- d_demo %>% 
  #filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id) %>%
  mutate(eth_group = ifelse(ethnicity=="White", "White", "Nonwhite"),
         ses_group = 
           ifelse(is.element(mom_ed, c("None", "Primary", "Some Secondary", "Secondary")), "low", "high"))
```

### Participants

```{r, echo=F}
sex_tab <- table(d_demo$sex)
ses_tab <- table(d_demo$ses_group)
eth_tab <- table(d_demo$ethnicity)
eth2_tab <- table(d_demo$eth_group)
momed_tab <- table(d_demo$mom_ed)
```

We analyze parent-reported Wordbank data from `r nrow(d_demo)` American English CDI: Words & Sentences administrations for children `r min(d_demo$age)` to `r max(d_demo$age)` months of age [@frank2017; @frank2021].
Full demographic data are not reported in some datasets contributed to Wordbank: sex was available for `r nrow(d_demo)-sum(is.na(d_demo$sex))` children, race/ethnicity was available for `r nrow(d_demo)-sum(is.na(d_demo$ethnicity))` children, and maternal education (a proxy for socioeconomic status; SES) was available for `r nrow(d_demo)-sum(is.na(d_demo$ses_group))` children.

The analysis of sex-based differences included CDI administrations from `r sex_tab["Female"]` female and `r sex_tab["Male"]` male children.
The analysis of race-based differences included data from `r eth_tab["White"]` white, `r eth_tab["Asian"]` Asian, `r eth_tab["Black"]` Black, `r eth_tab["Hispanic"]` Hispanic, and `r eth_tab["Other"]` "Other" children. 
Due to sparse data for many categories, we binarized participants' race/ethnicity as White (`r eth2_tab["White"]`) or Non-white (`r eth2_tab["Nonwhite"]`), recognizing that there may be important variation between groups that this will fail to capture. 
Data for the maternal education analysis included CDI data from children whose mother's had the following levels of education: `r momed_tab["Primary"]` with no more than primary school education, `r momed_tab["Some Secondary"]` with some secondary school, `r momed_tab["Secondary"]` with no more than secondary school, `r momed_tab["Some College"]` with some college, `r momed_tab["College"]` with no more than a college degree, `r momed_tab["Some Graduate"]` with no more than some graduate school, and `r momed_tab["Graduate"]` with a graduate degree.
Again due to data sparsity, we binarized the `r ses_tab["high"]` children whose mothers had at least some college or more as high maternal education (high-ME), and those whose mothers had at most high school (`r ses_tab["low"]` children) as low maternal education (low-ME).

## Rasch Model

The Rasch model, also known as the 1-parameter logistic (1PL) model, is the simplest Item Response Theory model, and is thus the easiest to use to investigate potential differences in item function across different groups of participants.
The Rasch model jointly estimates for each child $j$ a latent ability $\theta_j$, and for each item $i$ a difficulty parameter $b_i$.
In the model, the probability of child $j$ knowing (i.e., producing or understanding) a given item $i$ is

$$P_{i}(x_i = 1 | b_{i},\theta_j ) = \frac{1}{1 + e^{-D(\theta_j - b_i )}}$$

where $D$ is a constant scaling parameter ($D=1.702$) which makes the logistic closely match the ogive function in traditional factor analysis [@R-mirt; @reckase2009].
Child ability ($\theta$) and item difficulty ($b$) distributions are standardized (i.e., mean of 0), and expected to be normally-distributed. 
Children with high latent ability ($\theta$) will be more likely to produce any given item than children with lower latent ability, and more difficult items will be produced by fewer children (at any given $\theta$) than easier items.

In the multigroup Rasch model, an item's difficulty is allowed to vary by group. 
For example, in the sex-based multigroup model, item $i$'s difficulty is $b_{i}^{female}$ for females, and $b_{i}^{male}$ for males.
To identify DIF, a multigroup Rasch model will be fitted for each demographic dimension of interest (sex, maternal education, and ethnicity), and we will examine the between-group difficulty difference for each item (e.g., $d_i = b_{i}^{female} - b_{i}^{male}$).
If there is no DIF for a given item, then $d_i \approx 0$ as the two groups find the item equally difficult.

# Results

First, we will examine the size of demographic effects on language ability in a baseline Rasch model fitted without regard to demographic group.
Then we will fit a multigroup Rasch model for each demographic factor, and characterize the between-group differences in item difficulties.
Next, we will re-examine the size of demographic effects after pruning biased items from the CDI:WS using a varying threshold.
Finally, we measure the strength of association between sex-related differences in language input and the degree of sex-related DIF.

```{r load-data}
load(here("data/glimmer_prodWS_models.Rds"))
load(here("data/eng_Ws_prod_mod_1pl.Rds")) # m0h - baseline Rasch model, with coefs_1pl and fscores_1pl
#cor.test(d_demo$production, d_demo$ability) # .972
d_demo <- d_demo %>% left_join(fscores_1pl)

#d_demo %>% filter(!is.na(sex)) %>%
#  ggplot(aes(x=sex, y=ability)) +
#  geom_boxplot() + theme_classic()

sex_m <- d_demo %>% filter(!is.na(sex), !is.na(ability), age<26, age>22) %>% 
  group_by(sex) %>% summarise(ability=mean(ability),
                              production=mean(production),
                              n=n()) # 394 female, 408 male

ses_m <- d_demo %>% filter(!is.na(ses_group), !is.na(ability), age<26, age>22) %>% 
  group_by(ses_group) %>% summarise(ability=mean(ability),
                                    production=mean(production),
                                    n=n()) # 1014 high-SES, 140 low-SES

eth_m <- d_demo %>% filter(!is.na(eth_group), age<26, age>22) %>% 
  group_by(eth_group) %>% summarise(ability=mean(ability),
                                    production=mean(production),
                                    n=n()) # 144 nonwhite, 603 white

# sex_m[1,2:3] - sex_m[2,2:3] # average female vs. male advantage at 23-25 months of age: 0.78 theta, 69 words
# ses_m[1,2:3] - ses_m[2,2:3] # high vs. low advantage: 0.33 theta, 33 words
# eth_m[2,2:3] - eth_m[1,2:3] # white vs. nonwhite advantage: 0.52 theta, 45

# regression - report coefficients again after we cut some items?
d_demo$eth_group = factor(d_demo$eth_group, levels=c("Nonwhite","White"))
d_demo$sex = factor(d_demo$sex, levels=c("Male","Female"))
d_demo$ses_group = factor(d_demo$ses_group, levels=c("low","high"))
d_demo <- d_demo %>% mutate(age_sc = scale(age, scale=F))
eth_reg <- summary(lm(ability ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 white-, age*white+
ses_reg <- summary(lm(ability ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg <- summary(lm(ability ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

#ab_reg <- summary(lm(ability ~ age_sc * (eth_group + ses_group + sex), data=d_demo)) # R^2=.49
#prod_reg <- summary(lm(production ~ age_sc * (eth_group + ses_group + sex), data=d_demo)) # R^2=.51
```

## Demographic effects in baseline Rasch model
A baseline Rasch model was fitted to the entire dataset, without demographic information.
Figure 1 shows children's language ability vs. age by demographic group from the baseline Rasch model, which assumes no DIF (i.e., equal item parameters for all groups).
A linear regression for each demographic group, with age (centered) and its interaction, showed significant effects. 
Female children had higher language ability than male children ($\beta=0.56$, $p<.001$), with no significant interaction with age ($\beta=0.02$, $p=.10$).
High-SES children had higher language ability than low-SES children ($\beta=0.23$, $p=.02$), an advantage that grew with age ($\beta=0.11$, $p<.001$).
White children had higher language ability than non-white children ($\beta=0.50$, $p<.001$), an advantage that grew with age ($\beta=0.08$, $p<.001$).
We will re-examine these demographic regressions after trimming items showing extreme DIF.

<!-- Although the demographic language ability advantages go in the expected directions (i.e., favoring female, high-SES, and white children), note that magnitude of the differences vary somewhat across the age range, especially for SES and ethnicity.
Thus, although the IRT models are fitted to the entire age range, our comparisons of average ability will focus on the middle of the CDI:WS' intended age range, where there is great variation in the size and composition of children's vocabulary.
At ages 23-25 months, the average female vs. male advantage is an ability difference of $\theta_{f-m} = 0.78$, amounting to females knowing 69 more words, on average.
The average advantage for high-SES children at 23-25 months is $\theta_{high-low} = 0.33$, amounting to 33 more produced CDI items.
The average advantage for white vs. non-white children in this age range is $\theta_{w-nonw} = 0.52$, an average of 45 more produced CDI items.
-->

```{r baseline-ability-vs-age-hor, fig.env = "figure*", out.width="\\linewidth", fig.pos = "h", fig.align='center', fig.width=6.5, fig.height=2.8, set.cap.width=T, num.cols.cap=2, fig.cap = "Language ability vs. age by demographic group, from the baseline Rasch model."}
p1 <- d_demo %>% filter(!is.na(sex)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=sex)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") + 
  scale_color_discrete(type=sex_colors) +
  theme(legend.position = "top")

p2 <- d_demo %>% filter(!is.na(ses_group)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=ses_group)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") +
  scale_color_discrete(name = "SES", type=ses_colors) + 
  theme(legend.position = "top")

p3 <- d_demo %>% filter(!is.na(eth_group)) %>%
  ggplot(aes(x=jitter(age), y=ability, color=eth_group)) + 
  geom_point(alpha=.1) + geom_smooth() + theme_classic(base_size=9) + 
  xlab("Age (months)") + ylab("Language Ability") + 
  scale_color_discrete(name = "ethnicity", type=eth_colors) + 
  theme(legend.position = "top")

ggarrange(p1, p2, p3, ncol=3, nrow=1)
```

## Identifying biased CDI items

To aid in identifying CDI items with DIF we created GLIMMERs (Graphs of Logits Imputed Multiply with Means Equal; @stenhaug2021treading), which visualize between-item variation in group performance differences.
These parameters are drawn from a fitted multigroup Rasch model for each demographic variable (sex, SES, and race), with the assumption that the mean language ability in each group is the same (e.g. for sex, $\mu_{male}=\mu_{female}=0$), thus pushing all between-group variation into the item difficulty parameters.
For the case of sex, where we believe $\mu_{female}>\mu_{male}$, this means we may expect to find many items with difficulty $b_i^{female} < b_i^{male}$, but we can still examine the distribution of differences in item difficulty ($d_j = b_i^{male} - b_i^{female}$) for outliers.
GLIMMER plots show distributions of parameter differences rather than point estimates to convey the uncertainty about the existence of DIF.
These distributions are generated by drawing 10,000 imputations from the item parameter covariance matrix.

Figure 2 shows GLIMMERs for a selection of CDI items for sex (left), maternal education (middle), and race (right).
The full GLIMMERs, with all 680 CDI:WS items, are available on [OSF](https://osf.io/57rsw/?view_only=2b6ecb61fe08458293af7421d276932a), but were too large to include here.
It is important, however, to inspect the full plots, for if there is a cluster of items in a GLIMMER, the analyst may conclude that these items are strong candidates for DIF on that dimension.
For example, there is some clustering at the top and bottom of Figure 2: at the top, "vagina", "tights", "dress (object)", and "doll" form a cluster of items that are much more well-known for females, while at the bottom, "penis" stands out as much more well-known by males.
For the maternal education and race GLIMMERs (Figures 3 and 4)--and in the rest of the full sex GLIMMER, there are not clusters, but rather a continuum of smoothly varying differences with overlap. 
This makes identifying items with DIF quite difficult, as different methods are likely to yield inconsistent results [@stenhaug2021treading].
Hence, we next characterized the distributions of the item-level group difficulty differences, and measured the influence of pruning a varying number of items on the demographic effect sizes.

```{r glimmer-combo, fig.env = "figure*", out.width="\\linewidth", fig.pos = "h", fig.width=6.5, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the sex bias model. Words at the top are more well-known by females, while those at the bottom are more known for males."}
knitr::include_graphics("figs/smGLIMMER_combo.pdf")
```

```{r sex-glimmer, include=F, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.2, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the sex bias model. Words at the top are more well-known by females, while those at the bottom are more known for males."}
knitr::include_graphics("figs/smGLIMMER_sex_prodWS.pdf")
```

```{r ses-glimmer, include=F, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the SES bias model. Words at the top are more known by children in low-SES families, while those at the bottom are more known by those in high-SES families."}
knitr::include_graphics("figs/smGLIMMER_ses_prodWS.pdf")
```

```{r eth-glimmer, include=F, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "GLIMMER plot of a sample of CDI:WS words from the race bias model. Words at the top are more known by non-white children, while those at the bottom are more known by white children."}
knitr::include_graphics("figs/smGLIMMER_eth_prodWS.pdf")
```

```{r get-dif-dfs}
extract_group_df <- function(group_model, groups=c("Male","Female")) {
  Mit = as_tibble(coef(group_model, simplify=T)[[groups[1]]]$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)[[groups[1]]]$items),
           group1 = groups[1],
           group2 = groups[2]) %>%
    select(-g, -u) %>% 
    rename(d_g1 = d)
  Fit = as_tibble(coef(group_model, simplify=T)[[groups[2]]]$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)[[groups[2]]]$items)) %>%
    select(-g, -u) %>%
    rename(d_g2 = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_g2 - d_g1,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

get_extreme_item_difficulty_differences <- function(mm, sd_thresh=2, verbose=F) {
  #yfit <- dnorm(mm$d_diff, mean = mean(mm$d_diff), sd = sd(mm$d_diff)) 
  mm <- mm %>% mutate(d_diff = d_g2 - d_g1)
  max_dif = mean(mm$d_diff) + sd_thresh*sd(mm$d_diff)
  min_dif = mean(mm$d_diff) - sd_thresh*sd(mm$d_diff)
  mm <- mm %>% mutate(extreme = ifelse((d_diff > max_dif) | (d_diff < min_dif), T, F))
  if(verbose) print(paste("mininum difference:",round(min_dif,3), "maximum difference:",round(max_dif,3)))
  return(mm)
}

mm_sex <- extract_group_df(mod_intuitive_sex, groups=c("Male","Female"))
mm_ses <- extract_group_df(mod_intuitive_ses, groups=c("low","high"))
mm_eth <- extract_group_df(mod_intuitive_eth, groups=c("Nonwhite","White"))
```

```{r train-pruned-models, eval=F}
# varying SD thresholds (SLOW TO RUN)
prune_dat <- tibble()
for(thresh in seq(.25, 3.0, .25)) {
  tmp_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=thresh) # min: -.53 max: 1.06
  tmp_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=thresh) # min: -1.07 max: 1.13
  tmp_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=thresh) # min: -.64 max: 1.56
  prune_defs <- unique(c(subset(tmp_ses, extreme)$definition, 
                       subset(tmp_eth, extreme)$definition,
                       subset(tmp_sex, extreme)$definition))

  # prune the columns
  keep_cols <- which(!is.element(names(d_mat), prune_defs))
  #names(d_mat)[which(is.element(names(d_mat), prune_defs))]
  d_mat_pruned <- d_mat[,keep_cols]

  m0 <- mirt(d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(-data_id), 1, "Rasch", 
           technical=list(NCYCLES=1000))

  #coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  #  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
  fscores_1pl <- tibble(data_id = subset(d_mat_pruned, rowSums(d_mat_pruned)!=0)$data_id, 
                               ability_pr = fscores(m0, method = "MAP")[,1])

  d_demo_tmp <- d_demo %>% left_join(fscores_1pl) %>%
    mutate(age_sc = scale(age, center=T, scale=F))
  # join with d_demo, do the demo regressions again..

  eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo_tmp))$coefficients # R^2=.48 
  ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo_tmp))$coefficients # R^2=.56
  sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo_tmp))$coefficients # R^2=.59

  prune_dat = bind_rows(prune_dat, 
                        tibble(SDthreshold=thresh, 
                               extreme_sex=nrow(subset(tmp_sex, extreme)),
                               extreme_ses=nrow(subset(tmp_ses, extreme)),
                               extreme_eth=nrow(subset(tmp_eth, extreme)),
                               extreme_total = length(prune_defs),
                               sex_beta = sex_reg_pr[3,1],
                               ses_beta = ses_reg_pr[3,1],
                               eth_beta = eth_reg_pr[3,1]))
}

save(prune_dat, file=here("data/pruned_demo_coefs.Rdata"))
```

```{r}
# 2SD threshold
mm_sex <- get_extreme_item_difficulty_differences(mm_sex, sd_thresh=2.25) # min: -.53 max: 1.06
mm_ses <- get_extreme_item_difficulty_differences(mm_ses, sd_thresh=2.25) # min: -1.07 max: 1.13
mm_eth <- get_extreme_item_difficulty_differences(mm_eth, sd_thresh=2.25) # min: -.64 max: 1.56

all_extreme <- rbind(subset(mm_sex, extreme),
                     subset(mm_ses, extreme),
                     subset(mm_eth, extreme)) 
# length(unique(all_extreme$definition)) # 59
# sort(table(all_extreme$definition)) # candy, choo choo, daddy, duck, ... tractor, vroom

# overlaps extreme SES and race items:
# intersect(subset(mm_ses, extreme)$definition, subset(mm_eth, extreme)$definition)
# intersect(subset(mm_sex, extreme)$definition, subset(mm_eth, extreme)$definition)
# intersect(subset(mm_sex, extreme)$definition, subset(mm_ses, extreme)$definition)
#subset(all_extreme, group1=="Nonwhite")
```

```{r pruned-coefs-hor, fig.env = "figure*", out.width="\\linewidth", fig.pos = "h", fig.align='center', fig.width=6.5, fig.height=2.8, set.cap.width=T, num.cols.cap=2, fig.cap = "Size of demographic effects (regression coefficients) with different pruning thresholds."}
load(here("data/pruned_demo_coefs.Rdata"))
# horizontal version
pdat <- prune_dat %>% rename(c(beta_sex=sex_beta, beta_ses=ses_beta, beta_eth=eth_beta)) %>%
  pivot_longer(c("extreme_sex","extreme_ses","extreme_eth","beta_sex","beta_ses","beta_eth"), 
                           names_to=c("type", "demo"), names_sep="_") %>%
  pivot_wider(names_from=type, values_from=value) %>%
  mutate(Dimension = ifelse(demo=="ses", "maternal ed", ifelse(demo=="eth", "race", demo)))

pr1 <- pdat %>% filter(Dimension=="sex") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + #theme(legend.position = "bottom") + 
  labs(color="Excluded items") + ylim(-.7, -.4) + ggtitle("Sex") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

pr2 <- pdat %>% filter(Dimension=="maternal ed") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + 
  labs(color="Excluded items") + ylim(-.3, -.1) + ggtitle("Maternal Education") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

pr3 <- pdat %>% filter(Dimension=="race") %>%
  ggplot(aes(x=SDthreshold, y=beta, color=extreme)) + 
  geom_point() + 
  labs(color="Excluded items") + ylim(.35, .55) + ggtitle("Race") +
  xlab("Exclusion Threshold (SDs)") + ylab("Coefficient (beta)")

ggarrange(pr1, pr2, pr3, nrow=1, ncol=3, common.legend = T)
```

### Sex
For sex, the median difficulty difference (male-female) was `r round(median(mm_sex$d_diff),2)` (M=`r round(mean(mm_sex$d_diff),2)`, sd=`r round(sd(mm_sex$d_diff),2)`), with `r length(which(mm_sex$d_diff>0))`/680 items being easier for females than males.
The fact that the bulk of this distribution favors females shows that the female language advantage is pervasive across the CDI:WS, and suggests that it is likely to be a real ability advantage, unless the bulk of these items are actually spoken more often to female children than to male children.

### Maternal Education
For maternal education (ME), the median difficulty difference (low-high) was `r round(median(mm_ses$d_diff),2)` (M=`r round(mean(mm_ses$d_diff),2)`, sd=`r round(sd(mm_ses$d_diff),2)`), with `r length(which(mm_ses$d_diff>0))`/680 items being easier for high-ME than low-ME children. 
With the mean and median difficulty differences close to 0, and roughly half of the words favoring each ME group, it is tempting to conclude that the CDI:WS items are somewhat balanced with respect to ME (and thus SES, by proxy).

### Race
For race, the median difficulty difference (nonwhite-white) was `r round(median(mm_eth$d_diff),2)` (M=`r round(mean(mm_eth$d_diff),2)`, sd=`r round(sd(mm_eth$d_diff),2)`), with `r length(which(mm_eth$d_diff>0))`/680 items being easier for white than non-white children, revealing a fairly pervasive advantage for white children on CDI items.

<!--
## Recommended Pruning Threshold: 2SD
Considering just those items whose difficulty was >2SD from the mean difficulty difference identified 25 items with extreme sex-based difficulty differences, only 7 of which were easier for females, with the other 18 items favoring males.
For SES, 39 extrema were identified, only 13 of which were easier for low-SES children.
For race, 31 extrema were identified, only 11 of which were easier for non-white children.
Thus, while if anything the bulk of the extrema in the sex-based model advantaged the lower language ability group (males), the extrema in the other two models mostly favored the advantaged groups: 26/39 extreme items in the SES model favored high-SES children, and 20/31 items in the race model favored white children.
17 CDI items were identified as extrema in more than one model: "tractor" and "vroom" were extreme in all three models, "choo choo" was extreme in both SES and sex models, "give me five!" was extreme in both race and sex models, and 13 other items were extreme in both SES and race models ("grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy", "pet's name", "up", and "so").
In our final analysis, we explore the effect of pruning the 15 items that were extreme in both SES and race models.
-->

```{r items-to-remove}
prune_defs <- c("tractor", "vroom", "grrr", "moo", "quack quack", "uh oh", "duck", "owl", "candy", "gum", "walker", "daddy*", "pet's name", "up", "so") # also "choo choo" and "give me five!" ?
prune_cols <- which(!is.element(names(d_mat), prune_defs))
#names(d_mat)[which(is.element(names(d_mat), prune_defs))]
d_mat_pruned <- d_mat[,prune_cols]
```

```{r fit-pruned-model, eval=F}
m0 <- mirt(d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(-data_id), 1, "Rasch", 
           technical=list(NCYCLES=1000))

coefs_1pl <- as_tibble(coef(m0, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m0, simplify = TRUE)$items))
fscores_1pl <- tibble(data_id = d_mat_pruned %>% filter(rowSums(d_mat_pruned)!=0) %>% select(data_id), 
                             ability_pr = fscores(m0, method = "MAP")[,1])
fscores_1pl$data_id = fscores_1pl$data_id$data_id
save(file = here("data/eng_Ws_prod_mod_1pl_pruned.Rds"), "m0", "coefs_1pl", "fscores_1pl")
```

```{r, eval=F, include=F}
View(subset(mm_sex, extreme)) # 25 extreme (outside 2sd) - 7 favoring girls, 18 favoring boys
# possibly underestimating the female advantage?
View(subset(mm_ses, extreme)) # 39 extreme - 13 favoring low-SES, 26 favoring high-SES
View(subset(mm_eth, extreme)) # 31 extreme - 11 favoring nonwhite, 20 favoring white
# overestimating both high-SES and white advantages?
```

## Demographic effects after pruning

```{r, include=F}
# find minimal effect sizes on each demographich dimension
opt_sex = pdat[which(pdat$beta==max(subset(pdat, Dimension=="sex")$beta)),]$beta
opt_ME = pdat[which(pdat$beta==max(subset(pdat, Dimension=="maternal ed")$beta)),]$beta
opt_race = pdat[which(pdat$beta==min(subset(pdat, Dimension=="race")$beta)),]$beta

prune_dat <- prune_dat %>%
  mutate(SSE = (sex_beta-opt_sex)^2 + (ses_beta-opt_ME)^2 + (eth_beta-opt_race)^2,
         pct_overlap = extreme_total / (extreme_ses + extreme_eth + extreme_sex)) %>%
  arrange(SSE)
```

Given the smoothly-varying distributions of DIF shown by each demographic factor, we chose to evaluate multiple thresholds for pruning the extreme-valued items from each distribution.
For each demographic factor, we pruned items with difficulty difference from the mean at varying thresholds, from >.25 SD to >3 SD, in increments of .25 SD. 
At each SD threshold, a potentially different subset of items are excluded for each model, but with the goal of creating a single CDI that is less biased on all dimensions, we pruned the union of the subsets excluded from each model.
For example, pruning >2SD from the mean of each model excluded 25 sex-biased items, 39 ME-biased items, and 31 race-biased items, with their union being 76 unique items.
Figure 3 shows the demographic effects ($\beta$s) at different exclusion thresholds.

Ideally, we would find a single threshold that minimized the magnitude of coefficients (and thus bias) for all three demographics simultaneously.
Unfortunately, the effect size for each demographic variable was smallest (closest to 0) at different exclusion thresholds. 
The effect size of sex was smallest when almost all items were trimmed (>0.25 SD; $\beta_{sex} = -.47$; 661/680 total items trimmed; 433 due to sex extremity).
The effect size of maternal education (SES) was smallest when items >1.25SD in difficulty difference were trimmed ($\beta_{ME} = -.17$; 234 total items trimmed; 126 due to ME extremity). 
The effect size of race was minimized when items more extreme than 0.5 SD were trimmed ($\beta_{race} = .42$; 577 total items trimmed; 425 due to race extremity).

Where is the exclusion threshold that jointly minimizes the effect sizes of these demographic variables? 
The strict exclusion of all items with >.25 SD difficulty difference -- 661/680 (97%) of the CDI:WS -- best optimized this, but seems too extreme of a culling.
The next best thresholds are >.5 SD -- again, too extreme -- or 3 SD, a remarkably lax criterion that only excludes 19 items in total (11 due to sex bias).
Only slightly worse than these is >2.25 SD, which excludes 59 items (9% of the CDI:WS), and shows a modest effect of both race ($\beta=0.48$) and maternal education ($\beta=.19$), and a near-median effect size for sex ($\beta=-.59$).
We characterize these 59 potentially biased items below, and consider whether we might recommend pruning them from the CDI:WS.

## Characterizing biased items

Figure 4 shows the 59 items showing extreme bias (i.e., whose difficulty was >2.25 SD from the mean difficulty difference on one or more demographic dimension).
There were 22 items with extreme sex-based difficulty differences, only 7 of which were more known by females, with the other 15 items favoring males.
All but one of the sex-biased items are nouns, many of which are stereotypically associated with one gender more than the other, including stereotypically-male professions (e.g., "fireman").
For maternal education, 27 extrema were identified, only 10 of which were biased for low-SES children.
As for sex, most of the ME-biased items are nouns (15/27). 
Notably, many of the items more known by high-ME children were animals ("zebra", "owl") and animal sounds ("baa baa", "moo").
Items more known by low-ME children include a few sweet treats: "candy", "soda/pop", and "gum".
For race, 21 extrema were identified, only 8 of which were easier for non-white children.
The lexical class of the race-biased items was more varied: 10 were nouns, but many others were early-learned words and phrases (e.g., "up", "all gone", "uh-oh", "bye", "mommy").
Only 10 of the 59 items were extreme in more than one model, with most of the overlap being between race and ME (8 items).
Only 1 item ("vroom") was extreme for all three demographics.

```{r old-load-pruned, include=F, eval=F}
# (2sd threshold)
load(here("data/eng_Ws_prod_mod_1pl_pruned.Rds"))

d_demo <- d_demo %>% left_join(fscores_1pl)
# join with d_demo, do the demo regressions again..

eth_reg_pr <- summary(lm(ability_pr ~ age_sc * eth_group, data=d_demo))$coefficients # R^2=.48 
ses_reg_pr <- summary(lm(ability_pr ~ age_sc * ses_group, data=d_demo))$coefficients # R^2=.56
sex_reg_pr <- summary(lm(ability_pr ~ age_sc * sex, data=d_demo))$coefficients # R^2=.59

#white=.48
#highSES=0.20
#female=0.57

# table proportion of DIF items by lexical class -
# if due to environmental 
```

```{r demo-ability, include=F, out.width="\\linewidth", fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=7.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Scatterplots showing item difficulty per demographic group along each dimension."}

p1 <- mm_sex %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) + # color=extreme
  xlab("Male Difficulty") + ylab("Female Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

p2 <- mm_ses %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) +
  xlab("Low-ME Difficulty") + ylab("High-ME Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

p3 <- mm_eth %>%
  ggplot(aes(x=-d_g1, y=-d_g2)) + geom_point(alpha=.2) +
  xlab("Non-white Difficulty") + ylab("White Difficulty") +
  geom_text_repel(aes(label=definition)) +
  geom_abline(aes(slope=1, intercept=0), linetype="dashed")

# p1 <- mm_sex %>% 
#   ggplot(aes(x=d_diff)) + # , color=extreme, fill=extreme
#     geom_density(alpha=.2) + xlab("Ability")

ggarrange(p1, p2, p3, ncol=1)
```

```{r biased-words, include=T, out.width="\\linewidth", fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=3.5, fig.height=3.2, set.cap.width=T, num.cols.cap=1, fig.cap = "The 59 items showing extreme bias (difficulty difference $>2.25$ SD) for one or more demographic."}
knitr::include_graphics("figs/extrema2p25sd.pdf")
```

## Relating child-directed speech to demographic bias

Demographic differences in language ability are likely to be at least partially explained by differences in linguistic input received by children in different groups.
Indeed, input quantity (total daily tokens) and some measures of quality (e.g., lexical diversity: ratio of word types vs. tokens) have often been predictive of language learning outcomes in  demographic studies [@rowe2009differences; @huttenlocher1991early].
We used a corpus-based analysis to investigate the extent to which word frequency in child-directed speech to male vs. female children was predictive of the amount of sex-based bias shown by CDI items. 
Similar to the approach taken by @braginsky2016gender, we used the CHILDES corpus of transcripts from dyadic play sessions [@macwhinney2000childes], which are labeled with the sex of the target child, but not other demographic variables.
We found a total of 5213750 female-directed tokens and 6091950 male-directed tokens that matched 662 of the 680 CDI:WS items, normalized the word frequencies to tokens per million for each target sex, and calculated the log-odds of each word being spoken to a female (i.e., $log(p_{f} / p_{m}$), meaning unbiased words will have log-odds of 0, while those spoken more often to females will have log-odds $>0$, and those spoken more often males will have log-odds $<0$.
For example "doll" was spoken 156 times to females, and 138 times to males, thus $p_{f} = 156/(156+138) = .53$, $p_{m} = 1-p_{f} = 0.47$, and $log(p_{f}/p_{m})$ = 0.12), while "police" was spoken 138 times to females, and 225 times to males, and thus has log-odds = -0.49).
Overall, the correlation between the log-odds of a word being spoken to a female vs. a male child and the size of the female (vs. male) advantage for that CDI word was modest, but significant ($r=0.18$, $p<.001$), suggesting that some of the sex bias seen in CDI items is explained by differences in language input received by girls vs. boys.

```{r, include=F, out.width="\\linewidth", fig.env = "figure", fig.pos = "h", fig.align='center', fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "CHILDES frequency of input to male vs. female children."}
type_freqs = readRDS("../data/type_freqs.rds")
type_counts = readRDS("../data/type_counts.rds")
type_counts_smoothed = readRDS("../data/type_counts_smoothed.rds")

# total tokens:
type_counts %>% group_by(target_child_sex) %>% 
  summarise(total = sum(count))

type_counts_smoothed <- type_counts_smoothed %>%
  pivot_wider(id_cols=uni_lemma, names_from=sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male),
         FtoM = female / male)

# how biased are the extreme items we suggest removing in the corpus counts?
#sex_items_to_remove$definition

tmp <- left_join(mm_sex, type_counts_smoothed, by=c("definition"="uni_lemma"))

tmp %>% #filter(extreme==T) %>%
  ggplot(aes(x=propFemale, y=d_diff, color=extreme)) + 
  geom_point(alpha=.5) + theme_bw() +
  ylab("Female - Male Word Difficulty") +
  xlab("Proportion of Utterances to Females in CHILDES") +
  theme(legend.position = "bottom")

#subset(tmp, extreme==T & propFemale > .47 & propFemale < .53) # hose & dump

# subset(type_counts_smoothed, uni_lemma=="police")
#cor.test(log(tmp$FtoM), tmp$d_diff) # .18
#subset(type_counts_smoothed, uni_lemma=="police")
```

# Discussion

We investigated the CDI:WS, a popular parent-report measure of children's early vocabulary, for potential demographic bias, examining the distribution of words' estimated difficulties for children of mothers with high- vs. low-education children (a proxy for family SES), females vs. males, and for white vs. non-white children.
The IRT-based analysis revealed differential item functioning (DIF) for many items along each demographic dimension, but only in the case of sex were clear clusters of items that were more well-known to females (including feminine clothing and genitalia), and a clear item that was more well-known to males (male genitalia). 
For the rest of the items, and for SES- and race-based analysis, there was a smooth continuum of DIF, making the boundary of true DIF subjective, as this would rely on knowing the true difference in language ability between groups--which we reciprocally estimate from instruments like the CDI.
To move forward, we identified candidate DIF items systematically pruning the extremes of each distribution, excluding outliers for each demographic across a range of thresholds, and seeking to minimize the size of demographic effects.
Although many exclusion thresholds decreased the size of maternal education and race effects, the female advantage actually grew under most prunings, with the majority of excluded extrema being stereotypically-male nouns (e.g, "truck", "fireman").

This analysis highlights a fundamental difficulty of identifying DIF: an analyst must either know the expected magnitude of an ability difference between the two demographic groups, or know a set of "anchor" items that are equally difficult (i.e., unbiased) for both groups [for an overview, see @stenhaug2021treading].
But in the universe of children's early language, there is only a finite set of early-learned words to choose from--and we may expect many of them to be biased for various environmental reasons (e.g., boys and girls know their respective genitalia; children in Florida may not use mittens or skis).
Hence, the presence of DIF on a set of items could indicate a degree of varying linguistic input, rather than an actual advantage for one demographic over another.

Choosing an exclusion threshold that struck a balance of reducing race- and ME-based bias and not eliminating too many items, we identified 59 suspect items--the majority of which favored white/high-ME groups.
These extrema were predominantly nouns, including many animals (esp. for maternal education), vehicles and professions (esp. for sex). 
Notably, no verbs and few syntactically complex items were identified as outliers, which may suggest that these lexical classes are less prone to environmental influence, and perhaps more generally unbiased. <!--or something..seems important, as nouns are more referential/context-dependent-->
Pruning these 59 outliers would decrease the demographic effects of maternal education (unpruned $\beta=.23$; pruned $\beta=.19$) and race (unpruned $\beta=.50$; pruned $\beta=.48$).
In contrast, the majority of the sex-biased outliers were easier for the disadvantaged group (males), meaning that the removal of these extrema increased the female language advantage on the CDI:WS (unpruned $\beta=0.56$; pruned $\beta=.59$).
In aggregate, these analyses seem to suggest that the CDI:WS may slightly overestimate the magnitude of differences in language ability along the dimensions of maternal education and race, and perhaps underestimate the female advantage.

This investigation is only a first step in measuring demographic bias in the items on the CDI:WS.
More research is needed to determine which CDI items that show DIF are more subject to environmental influence, as we found only a modest association between the relative frequency of words in child-directed speech to boys vs. girls and the amount of sex-based DIF.
Examining the relative frequency of CDI words across households with varying maternal education and race may reveal similarly weak associations, or there may be strong associations only among particular categories of words.
It may be that particular activity contexts--popular with some demographic groups, and not others--may be predictive. 
For example, many of the extrema favoring high-ME children are animals and animal sounds (e.g., "quack quack", "woof woof", "baa baa", "duck", "sheep", "giraffe", "zebra"): do high-SES households visit the zoo more often, or perhaps engage in other activities related to naming animals and noises they make (e.g., animal noise toys)?
If certain activities are driving early word learning for high-ME children, what are the activities (and associated vocabulary) that low-SES households are instead engaging in?
Note, too, that SES, race, and sex are only three of many possible demographic dimensions of interest. For example, geographic region is likely predictive of children's early experience with (and thus knowledge of) CDI items related to winter weather (e.g., "snow", "mittens") or outdoors activities (e.g. "camping").
A truly fair test of children's early vocabulary would contain a representative and balanced sample of words from all activities that children engage in, across demographic groups.

<!--However, the best way to estimate endogenous differences in ability (rather than exogenous differences, e.g. in the language environment) would be to conduct controlled, in-lab experiments measuring children's ability to learn novel words.-->

<!-- For example, let male language ability be drawn from a standard normal $\mu_{male} ~ N(0,1)$, with female language ability slightly higher, on average ($\mu_{female} = \mu_{male} + 0.1$).
Then we would expect items to be an average of $0.1$ easier for females than for males, and we might identify items that are instead easier for males than females as showing undesirable DIF.
And yet, without knowing the actual ability difference between two demographic groups--for which we also rely upon our tests, it is difficult to adjudicate which items show DIF, and which do not.
-->

# Acknowledgements

[Redacted for anonymous review.]
<!--We thank members of the Language and Cognition lab for their feedback.-->

# References

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent